### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

---
title: "gtsummary"
author: "Daniel D. Sjoberg, Karissa Whiting, Margaret Hannum, Emily C. Zabor, Michael Curry"
output:
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(gtsummary)
library(tidyverse)
library(gt)

theme_gtsummary_compact()
```

## Introduction

## Data Summaries

The most common summary needed for research projects are simple summaries of data sets.
To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
The data set has label attributes (using the labelled package) for column names.

```{r}
trial %>%
  imap_dfr(
    ~tibble(
      colname = .y, 
      label = attr(.x, "label"), 
      class = class(.x), 
      values = trial %>% select(.y) %>% distinct() %>% 
        arrange(!!rlang::sym(.y)) %>% slice(1:4) %>% pull(.y) %>% 
        na.omit() %>% {paste("`", ., "`", collapse = ", ")}
    ) %>%
      mutate(
        values = ifelse(length(unique(na.omit(trial[[.y]]))) > 4,
                        paste(values, ", ..."),
                        values) 
      )
  ) %>%
  knitr::kable(caption = "Example data frame, `trial`")
```

### `tbl_summary()`

The `tbl_summary()` function can be used to easily create a basic summary statistic table.
This is often the first table of clinical manuscripts and describes characteristics of the cohort under study.
A simple example is shown below with some basic customizations using the function’s main arguments.
Notably, by specifying the “by” argument, you can divide your summary statistic into comparator groups.
In this case, we will split the table by treatment arms.

```{r, echo=TRUE}
tbl_summary_1 <-
  trial %>%
  select(age, grade, response, trt) %>%
  tbl_summary(by = trt)
```

```{r, echo=FALSE, include = FALSE}
tbl_summary_1 %>% 
  as_gt() %>%
  gtsave(file = "tbl_summary_1.png")
```

\begin{figure}[h!]
  \caption{Simple `tbl\_summary()` example}
  \label{fig:summary_basic}
  \includegraphics[height=5cm]{tbl_summary_1.png}
  \centering
\end{figure}

Figure \ref{fig:summary_basic} is the most bsaic table created with `tbl_summary()`

### `tbl_svysummary()`

### `tbl_cross()`

### `tbl_survfit()`

### Customization

## Model Summaries

### `tbl_regression()`

### `tbl_uvregression()`

## In-line Reporting

## Merging and Stacking

## Themes

## Print Engines

### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

@@
 theme_gtsummary_compact()
+set_gtsummary_theme(list(
+  "as_gt-lst:addl_cmds" = list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino")))
+))
 ```
 
-## Introduction
+# Introduction
 
-## Data Summaries
-
-The most common summary needed for research projects are simple summaries of data sets.
-To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
-The data set has label attributes (using the labelled package) for column names.
+# Data Summaries
 
@@
 
-### `tbl_summary()`
-
-The `tbl_summary()` function can be used to easily create a basic summary statistic table.
-This is often the first table of clinical manuscripts and describes characteristics of the cohort under study.
-A simple example is shown below with some basic customizations using the function’s main arguments.
-Notably, by specifying the “by” argument, you can divide your summary statistic into comparator groups.
-In this case, we will split the table by treatment arms.
+## `tbl_summary()`
 
@@
   as_gt() %>%
-  gtsave(file = "tbl_summary_1.png")
+  gtsave(file = "summary_basic.png")
+```
+
+```{r}
+tibble::tribble(
+  ~Argument,       ~Description,
+  "`label`",       "specify the variable labels printed in table",
+  "`type`",        "specify the variable type (e.g. continuous, categorical, etc.)",
+  "`statistic`",   "change the summary statistics presented",
+  "`digits`",      "number of digits the summary statistics will be rounded to",
+  "`missing`",     "whether to display a row with the number of missing observations",
+  "`missing_text`","text label for the missing number row",
+  "`sort`",        "change the sorting of categorical levels by frequency",
+  "`percent`",     "print column, row, or cell percentages"
+) %>%
+  knitr::kable(caption = "`tbl_summary()` function arguments") 
+```
+
+```{r, echo=TRUE}
+tbl_summary_2 <-
+  trial %>%
+  select(age, grade, response, trt) %>%
+  tbl_summary(by = trt,
+              label = age ~ "Patient Age",
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} / {N} ({p}%)"),
+              digits = list(age ~ 2, 
+                            all_categorical() ~ c(0, 0, 1)),
+              missing = "no")
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_summary_2 %>% 
+  as_gt() %>%
+  gtsave(file = "summary_plus.png")
+```
+
+```{r}
+tibble::tribble(
+  ~Function,             ~Description,
+  "`add_p()`",           "add p-values to the output comparing values across groups",   
+  "`add_overall()`",     "add a column with overall summary statistics",   
+  "`add_n()`",           "add a column with N (or N missing) for each variable",
+  "`add_difference()`",  "add column for difference between two group, confidemce interval, and p-value",
+  "`add_stat_label()`",  "add label for the summary statistics shown in each row",   
+  "`add_stat()`",        "generic function to add a column with user-defined values",
+  "`add_q()`",           "add a column of q values to control for multiple comparisons"   
+) %>%
+  knitr::kable(caption = "`tbl_summary()` function arguments") 
 ```
@@
   \label{fig:summary_basic}
-  \includegraphics[height=5cm]{tbl_summary_1.png}
+  \includegraphics[height=5cm]{summary_basic.png}
   \centering
@@
 
-Figure \ref{fig:summary_basic} is the most bsaic table created with `tbl_summary()`
+## `tbl_svysummary()`
 
-### `tbl_svysummary()`
+## `tbl_cross()`
 
-### `tbl_cross()`
+## `tbl_survfit()`
 
-### `tbl_survfit()`
+## Customization
 
-### Customization
+# Model Summaries
 
-## Model Summaries
+## `tbl_regression()`
 
-### `tbl_regression()`
+## `tbl_uvregression()`
 
-### `tbl_uvregression()`
+# In-line Reporting
 
-## In-line Reporting
+# Merging and Stacking
 
-## Merging and Stacking
+# Themes
 
-## Themes
-
-## Print Engines
+# Print Engines

### 2020-12-09 tbl_tbl_summary_args.tex
in progress

Table: `tbl_summary()` function arguments

|Argument       |Description                                                      |
|:--------------|:----------------------------------------------------------------|
|`label`        |specify the variable labels printed in table                     |
|`type`         |specify the variable type (e.g. continuous, categorical, etc.)   |
|`statistic`    |change the summary statistics presented                          |
|`digits`       |number of digits the summary statistics will be rounded to       |
|`missing`      |whether to display a row with the number of missing observations |
|`missing_text` |text label for the missing number row                            |
|`sort`         |change the sorting of categorical levels by frequency            |
|`percent`      |print column, row, or cell percentages                           |

### 2020-12-09 tbl_trial_desc.tex
in progress

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Example data frame, `trial`}
\centering
\begin{tabular}[t]{l|l|l|l}
\hline
colname & label & class & values\\
\hline
trt & Chemotherapy Treatment & character & ` Drug A `, ` Drug B `\\
\hline
age & Age & numeric & ` 6 `, ` 9 `, ` 10 `, ` 17 ` , ...\\
\hline
marker & Marker Level (ng/mL) & numeric & ` 0.003 `, ` 0.005 `, ` 0.013 `, ` 0.015 ` , ...\\
\hline
stage & T Stage & factor & ` T1 `, ` T2 `, ` T3 `, ` T4 `\\
\hline
grade & Grade & factor & ` I `, ` II `, ` III `\\
\hline
response & Tumor Response & integer & ` 0 `, ` 1 `\\
\hline
death & Patient Died & integer & ` 0 `, ` 1 `\\
\hline
ttdeath & Months to Death/Censor & numeric & ` 3.53 `, ` 5.33 `, ` 6.32 `, ` 7.27 ` , ...\\
\hline
\end{tabular}
\end{table}

### 2020-12-09 rmarkdown-gtsummary.Rmd
Update rmarkdown-gtsummary.Rmd

@@
 ```{r}
+tbl_trial_desc <- 
   trial %>%
@@
       )
-  ) %>%
+  )
+
+tbl_trial_desc %>%
   knitr::kable(caption = "Example data frame, `trial`")
+
+tbl_trial_desc %>%
+  gt(caption = "Example data frame, `trial`") %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  cat() %>%
+  readr::write_lines("tbl_trial_desc.tex")
 ```

### 2020-12-09 RJtemplate_gtsummary_submission.tex

% !TeX root = RJwrapper.tex
\title{Reproducible Summary Tables with gtsummary}
\author{by Daniel D. Sjoberg, Karissa Whiting, Margie Hannum, Michael Curry}

\maketitle

\abstract{
An abstract of less than 150 words.
}

\section{Section title in sentence case}

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

\section{Another section}

This section may contain a figure such as Figure~\ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo-5}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

\section{Another section}

There will likely be several sections, perhaps including code snippets, such as:

\begin{example}
  x <- 1:10
  result <- myFunction(x)
\end{example}

\section{Summary}

This file is only a basic article template. For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{https://journal.r-project.org/share/author-guide.pdf}{Instructions for Authors}.

\bibliography{RJreferences}

\address{Author One\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author1@work}}

\address{Author Two\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author2@work}}

\address{Author Three\\
  Affiliation\\
  Address\\
  Country\\
  (ORCiD if desired)\\
  \email{author3@work}}

### 2020-12-09 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 
-\section{Section title in sentence case}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+\hypertarget{data-summaries}{%
+\section{Data Summaries}\label{data-summaries}}
 
-\section{Another section}
+\include{tbl_trial_desc}
 
-This section may contain a figure such as Figure~\ref{figure:rlogo}.
+\section{test}
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics{Rlogo-5}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
-\end{figure}
-
-\section{Another section}
-
-There will likely be several sections, perhaps including code snippets, such as:
-
-\begin{example}
-  x <- 1:10
-  result <- myFunction(x)
-\end{example}
-
-\section{Summary}
-
-This file is only a basic article template.
-For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{https://journal.r-project.org/share/author-guide.pdf}{Instructions for Authors}.
+\begin{longtable}[]{@{}llll@{}}
+\caption{Example data frame, \texttt{trial}}\tabularnewline
+\toprule
+colname & label & class & values\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+colname & label & class & values\tabularnewline
+\midrule
+\endhead
+trt & Chemotherapy Treatment & character & \texttt{Drug\ A},
+\texttt{Drug\ B}\tabularnewline
+age & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17} ,
+\ldots{}\tabularnewline
+marker & Marker Level (ng/mL) & numeric & \texttt{0.003},
+\texttt{0.005}, \texttt{0.013}, \texttt{0.015} , \ldots{}\tabularnewline
+stage & T Stage & factor & \texttt{T1}, \texttt{T2}, \texttt{T3},
+\texttt{T4}\tabularnewline
+grade & Grade & factor & \texttt{I}, \texttt{II},
+\texttt{III}\tabularnewline
+response & Tumor Response & integer & \texttt{0},
+\texttt{1}\tabularnewline
+death & Patient Died & integer & \texttt{0}, \texttt{1}\tabularnewline
+ttdeath & Months to Death/Censor & numeric & \texttt{3.53},
+\texttt{5.33}, \texttt{6.32}, \texttt{7.27} , \ldots{}\tabularnewline
+\bottomrule
+\end{longtable}
 

### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

--- 2020-12-09 rmarkdown-gtsummary.Rmd
@@
     ~tibble(
-      colname = .y, 
+      colname = str_glue('`"{.y}"`'), 
       label = attr(.x, "label"), 
@@
   ) %>%
-  knitr::kable(caption = "Example data frame, `trial`") 
+  gt(caption = "Example data frame, `trial`") %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  readr::write_lines("tbl_trial_desc.tex")
 ```
@@
 )  %>%
-  knitr::kable(caption = "`tbl_summary()` function arguments") 
+  gt(caption = "`tbl_summary()` function arguments") %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  readr::write_lines("tbl_tbl_summary_args.tex")
+
 ```

### 2020-12-09 tbl_tbl_summary_args.tex
in progress

@@
-Table: `tbl_summary()` function arguments
-
-|Argument       |Description                                                      |
-|:--------------|:----------------------------------------------------------------|
-|`label`        |specify the variable labels printed in table                     |
-|`type`         |specify the variable type (e.g. continuous, categorical, etc.)   |
-|`statistic`    |change the summary statistics presented                          |
-|`digits`       |number of digits the summary statistics will be rounded to       |
-|`missing`      |whether to display a row with the number of missing observations |
-|`missing_text` |text label for the missing number row                            |
-|`sort`         |change the sorting of categorical levels by frequency            |
-|`percent`      |print column, row, or cell percentages                           |
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{ll}
+\toprule
+Argument & Description \\ 
+\midrule
+\texttt{label} & specify the variable labels printed in table \\ 
+\texttt{type} & specify the variable type (e.g. continuous, categorical, etc.) \\ 
+\texttt{statistic} & change the summary statistics presented \\ 
+\texttt{digits} & number of digits the summary statistics will be rounded to \\ 
+\texttt{missing} & whether to display a row with the number of missing observations \\ 
+\texttt{missing\_text} & text label for the missing number row \\ 
+\texttt{sort} & change the sorting of categorical levels by frequency \\ 
+\texttt{percent} & print column, row, or cell percentages \\ 
+\bottomrule
+\end{longtable}

### 2020-12-09 tbl_trial_desc.tex
in progress

@@
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-1}Example data frame, `trial`}
-\centering
-\begin{tabular}[t]{l|l|l|l}
-\hline
-colname & label & class & values\\
-\hline
-trt & Chemotherapy Treatment & character & ` Drug A `, ` Drug B `\\
-\hline
-age & Age & numeric & ` 6 `, ` 9 `, ` 10 `, ` 17 ` , ...\\
-\hline
-marker & Marker Level (ng/mL) & numeric & ` 0.003 `, ` 0.005 `, ` 0.013 `, ` 0.015 ` , ...\\
-\hline
-stage & T Stage & factor & ` T1 `, ` T2 `, ` T3 `, ` T4 `\\
-\hline
-grade & Grade & factor & ` I `, ` II `, ` III `\\
-\hline
-response & Tumor Response & integer & ` 0 `, ` 1 `\\
-\hline
-death & Patient Died & integer & ` 0 `, ` 1 `\\
-\hline
-ttdeath & Months to Death/Censor & numeric & ` 3.53 `, ` 5.33 `, ` 6.32 `, ` 7.27 ` , ...\\
-\hline
-\end{tabular}
-\end{table}
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{clll}
+\toprule
+colname & label & class & values \\ 
+\midrule
+\texttt{\textquotedbl{}trt\textquotedbl{}} & Chemotherapy Treatment & character & \texttt{Drug A}, \texttt{Drug B} \\ 
+\texttt{\textquotedbl{}age\textquotedbl{}} & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17} , ... \\ 
+\texttt{\textquotedbl{}marker\textquotedbl{}} & Marker Level (ng/mL) & numeric & \texttt{0.003}, \texttt{0.005}, \texttt{0.013}, \texttt{0.015} , ... \\ 
+\texttt{\textquotedbl{}stage\textquotedbl{}} & T Stage & factor & \texttt{T1}, \texttt{T2}, \texttt{T3}, \texttt{T4} \\ 
+\texttt{\textquotedbl{}grade\textquotedbl{}} & Grade & factor & \texttt{I}, \texttt{II}, \texttt{III} \\ 
+\texttt{\textquotedbl{}response\textquotedbl{}} & Tumor Response & integer & \texttt{0}, \texttt{1} \\ 
+\texttt{\textquotedbl{}death\textquotedbl{}} & Patient Died & integer & \texttt{0}, \texttt{1} \\ 
+\texttt{\textquotedbl{}ttdeath\textquotedbl{}} & Months to Death/Censor & numeric & \texttt{3.53}, \texttt{5.33}, \texttt{6.32}, \texttt{7.27} , ... \\ 
+\bottomrule
+\end{longtable}

### 2020-12-09 RJtemplate_gtsummary_submission.tex

--- 2020-12-09 RJtemplate_gtsummary_submission.tex
@@
 
-\section{Section title in sentence case}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+\hypertarget{data-summaries}{%
+\section{Data Summaries}\label{data-summaries}}
 
-\section{Another section}
-
-This section may contain a figure such as Figure~\ref{figure:rlogo}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics{Rlogo-5}
-  \caption{The logo of R.}
-  \label{figure:rlogo}
-\end{figure}
-
-\section{Another section}
-
-There will likely be several sections, perhaps including code snippets, such as:
-
-\begin{example}
-  x <- 1:10
-  result <- myFunction(x)
-\end{example}
-
-\section{Summary}
-
-This file is only a basic article template.
-For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{https://journal.r-project.org/share/author-guide.pdf}{Instructions for Authors}.
+% adding table describing trial dataset
+\input{tbl_trial_desc.tex}
 

### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

--- 2020-12-09 rmarkdown-gtsummary.Rmd
@@
   ) %>%
-  knitr::kable(caption = "Example data frame, `trial`") 
+  gt() %>%
+  fmt_markdown(everything()) %>%
+  cols_align("left", everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}Example data frame, \\texttt{trial}}\\\\\n\\toprule")) %>%
+  readr::write_lines("tbl_trial_desc.tex")
 ```
@@
 ) %>%
-  knitr::kable(caption = "`tbl_summary()` function arguments") 
+  gt() %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl_summary()} function arguments}\\\\\n\\toprule")) %>%
+  readr::write_lines("tbl_tbl_summary_args.tex")
+
 ```
@@
   "`add_q()`",           "add a column of q values to control for multiple comparisons"   
-) %>%
-  knitr::kable(caption = "`tbl_summary()` function arguments") 
+)%>%
+  gt(caption = "`tbl_summary()` functions to add information") %>%
+  fmt_markdown(everything()) %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl_summary()} functions to add information}\\\\\n\\toprule")) %>%
+  readr::write_lines("tbl_tbl_summary_family.tex")
 ```

### 2020-12-09 tbl_tbl_summary_family.tex
in progress

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{ll}
\caption{\label{tab:}\texttt{tbl_summary()} functions to add information}\\
\toprule
Function & Description \\ 
\midrule
\texttt{add\_p()} & add p-values to the output comparing values across groups \\ 
\texttt{add\_overall()} & add a column with overall summary statistics \\ 
\texttt{add\_n()} & add a column with N (or N missing) for each variable \\ 
\texttt{add\_difference()} & add column for difference between two group, confidemce interval, and p-value \\ 
\texttt{add\_stat\_label()} & add label for the summary statistics shown in each row \\ 
\texttt{add\_stat()} & generic function to add a column with user-defined values \\ 
\texttt{add\_q()} & add a column of q values to control for multiple comparisons \\ 
\bottomrule
\end{longtable}

### 2020-12-09 tbl_trial_desc.tex
Update tbl_trial_desc.tex
in progress

--- 2020-12-09 tbl_trial_desc.tex
@@
-\begin{table}
-
-\caption{\label{tab:unnamed-chunk-1}Example data frame, `trial`}
-\centering
-\begin{tabular}[t]{l|l|l|l}
-\hline
-colname & label & class & values\\
-\hline
-trt & Chemotherapy Treatment & character & ` Drug A `, ` Drug B `\\
-\hline
-age & Age & numeric & ` 6 `, ` 9 `, ` 10 `, ` 17 ` , ...\\
-\hline
-marker & Marker Level (ng/mL) & numeric & ` 0.003 `, ` 0.005 `, ` 0.013 `, ` 0.015 ` , ...\\
-\hline
-stage & T Stage & factor & ` T1 `, ` T2 `, ` T3 `, ` T4 `\\
-\hline
-grade & Grade & factor & ` I `, ` II `, ` III `\\
-\hline
-response & Tumor Response & integer & ` 0 `, ` 1 `\\
-\hline
-death & Patient Died & integer & ` 0 `, ` 1 `\\
-\hline
-ttdeath & Months to Death/Censor & numeric & ` 3.53 `, ` 5.33 `, ` 6.32 `, ` 7.27 ` , ...\\
-\hline
-\end{tabular}
-\end{table}
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{ll}
+\caption{\label{tab:}tbl_summary() function arguments}\\
+\toprule
+Argument & Description \\ 
+\midrule
+\texttt{label} & specify the variable labels printed in table \\ 
+\texttt{type} & specify the variable type (e.g. continuous, categorical, etc.) \\ 
+\texttt{statistic} & change the summary statistics presented \\ 
+\texttt{digits} & number of digits the summary statistics will be rounded to \\ 
+\texttt{missing} & whether to display a row with the number of missing observations \\ 
+\texttt{missing\_text} & text label for the missing number row \\ 
+\texttt{sort} & change the sorting of categorical levels by frequency \\ 
+\texttt{percent} & print column, row, or cell percentages \\ 
+\bottomrule
+\end{longtable}

### 2020-12-09 RJtemplate_gtsummary_submission.tex

@@
 \input{tbl_trial_desc.tex}
+
+\hypertarget{tbl_summary}{%
+\subsection{\texorpdfstring{\texttt{tbl\_summary()}}{tbl\_summary()}}\label{tbl_summary}}
+
+% code for basic tbl_summary()
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{tbl\_summary\_}\DecValTok{1}\NormalTok{ <{-}}
+\StringTok{  }\NormalTok{trial }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(age, grade, response, trt) }\OperatorTok{\%>\%}
+\StringTok{  }\KeywordTok{tbl\_summary}\NormalTok{(}\DataTypeTok{by =}\NormalTok{ trt)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{tbl_svysummary}{%
+\subsection{\texorpdfstring{\texttt{tbl\_svysummary()}}{tbl\_svysummary()}}\label{tbl_svysummary}}
+
+\hypertarget{tbl_cross}{%
+\subsection{\texorpdfstring{\texttt{tbl\_cross()}}{tbl\_cross()}}\label{tbl_cross}}
+
+\hypertarget{tbl_survfit}{%
+\subsection{\texorpdfstring{\texttt{tbl\_survfit()}}{tbl\_survfit()}}\label{tbl_survfit}}
+
+\hypertarget{customization}{%
+\subsection{Customization}\label{customization}}
+
+\hypertarget{model-summaries}{%
+\section{Model Summaries}\label{model-summaries}}
+
+\hypertarget{tbl_regression}{%
+\subsection{\texorpdfstring{\texttt{tbl\_regression()}}{tbl\_regression()}}\label{tbl_regression}}
+
+\hypertarget{tbl_uvregression}{%
+\subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}\label{tbl_uvregression}}
+
+\hypertarget{in-line-reporting}{%
+\section{In-line Reporting}\label{in-line-reporting}}
+
+\hypertarget{merging-and-stacking}{%
+\section{Merging and Stacking}\label{merging-and-stacking}}
+
+\hypertarget{themes}{%
+\section{Themes}\label{themes}}
+
+\hypertarget{print-engines}{%
+\section{Print Engines}\label{print-engines}}
 

### 2020-12-09 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

--- 2020-12-09 RJtemplate_gtsummary_submission.tex
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}
 
-\hypertarget{data-summaries}{%
-\section{Data Summaries}\label{data-summaries}}
+\section{Data Summaries}
 
@@
 \input{tbl_trial_desc.tex}
+
+\subsection{\texorpdfstring{\texttt{tbl\_summary()}}{tbl\_summary()}}
+
+% code for basic tbl_summary()
+\begin{example}
+tbl_summary_1 <-
+  trial %>%
+  select(age, grade, response, trt) %>%
+  tbl_summary(by = trt)
+\end{example}
+
+\subsection{\texorpdfstring{\texttt{tbl\_svysummary()}}{tbl\_svysummary()}}
+
+\subsection{\texorpdfstring{\texttt{tbl\_cross()}}{tbl\_cross()}}
+
+\subsection{\texorpdfstring{\texttt{tbl\_survfit()}}{tbl\_survfit()}}
+
+\subsection{Customization}
+
+\section{Model Summaries}
+
+\subsection{\texorpdfstring{\texttt{tbl\_regression()}}{tbl\_regression()}}
+
+\hypertarget{tbl_uvregression}{%
+\subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}\label{tbl_uvregression}}
+
+\hypertarget{in-line-reporting}{%
+\section{In-line Reporting}\label{in-line-reporting}}
+
+\hypertarget{merging-and-stacking}{%
+\section{Merging and Stacking}\label{merging-and-stacking}}
+
+\hypertarget{themes}{%
+\section{Themes}\label{themes}}
+
+\hypertarget{print-engines}{%
+\section{Print Engines}\label{print-engines}}
 

### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

@@
   as.character() %>%
-  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl_summary()} function arguments}\\\\\n\\toprule")) %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl\\_summary()} function arguments}\\\\\n\\toprule")) %>%
   readr::write_lines("tbl_tbl_summary_args.tex")
@@
   as.character() %>%
-  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl_summary()} functions to add information}\\\\\n\\toprule")) %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl\\_summary()} functions to add information}\\\\\n\\toprule")) %>%
   readr::write_lines("tbl_tbl_summary_family.tex")

### 2020-12-09 rmarkdown-gtsummary.Rmd
i nprogress

@@
   select(age, grade, response, trt) %>%
-  tbl_summary(by = trt,
+  tbl_summary(
+    by = trt,
     label = age ~ "Patient Age",
@@
                   all_categorical() ~ c(0, 0, 1)),
-              missing = "no")
+    missing = "no"
+  )
 ```
@@
   readr::write_lines("tbl_tbl_summary_family.tex")
+```
+
+```{r, echo=TRUE}
+tbl_summary_3 <-
+  trial %>%
+  select(age, grade, response, trt) %>%
+  tbl_summary(by = trt, missing = "no") %>%
+  add_p(test = all_continuous() ~ "t.test",
+        pvalue_fun = ~style_pvalue(., digits = 2)) %>%
+  add_n()
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_summary_3 %>% 
+  as_gt() %>%
+  gtsave(file = "summary_plus_plus.png")
 ```

### 2020-12-09 rmarkdown-gtsummary.Rmd
in progress

@@
     by = trt,
+    type = all_continuous() ~ "continuous2",
     label = age ~ "Patient Age",
-    statistic = list(all_continuous() ~ "{mean} ({sd})",
+    statistic = list(all_continuous() ~ c("{N_nonmiss}", 
+                                          "{mean} ({sd})", 
+                                          "{median} ({p25}, {p75})", 
+                                          "{min}, {max}"),
                      all_categorical() ~ "{n} / {N} ({p}%)"),
-    digits = list(age ~ 2, 
-                  all_categorical() ~ c(0, 0, 1)),
+    digits = all_categorical() ~ c(0, 0, 1),
     missing = "no"

### 2020-12-09 RJtemplate_gtsummary_submission.tex

@@
 \begin{example}
-tbl_summary_1 <-
 trial %>%
@@
 \end{example}
+\begin{figure}[h!]
+  \includegraphics[height=5cm]{summary_basic.png}
+  \centering
+\end{figure}
+
+% code for tbl_summary() with arguments
+\input{tbl_tbl_summary_args.tex}
+\begin{example}
+trial %>%
+  select(age, grade, response, trt) %>%
+  tbl_summary(
+    by = trt,
+    type = all_continuous() ~ "continuous2",
+    label = age ~ "Patient Age",
+    statistic = list(all_continuous() ~ c("{N_nonmiss}", 
+                                          "{mean} ({sd})", 
+                                          "{median} ({p25}, {p75})", 
+                                          "{min}, {max}"),
+                     all_categorical() ~ "{n} / {N} ({p}%)"),
+    digits = all_categorical() ~ c(0, 0, 1),
+    missing = "no"
+  )
+\end{example}
+\begin{figure}[h!]
+  \includegraphics[height=5cm]{summary_plus.png}
+  \centering
+\end{figure}
+
+% code for tbl_summary() with arguments
+\input{tbl_tbl_summary_family.tex}
+\begin{example}
+trial %>%
+  select(age, grade, response, trt) %>%
+  tbl_summary(by = trt, missing = "no") %>%
+  add_p(test = all_continuous() ~ "t.test",
+        pvalue_fun = ~style_pvalue(., digits = 2)) %>%
+  add_n()
+\end{example}
+\begin{figure}[h!]
+  \includegraphics[height=5cm]{summary_plus_plus.png}
+  \centering
+\end{figure}
 

### 2020-12-10 rmarkdown-gtsummary.Rmd
in progress

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 library(gtsummary)
@@
 
-```{r}
+```{r, results='asis'}
 trial %>%
@@
 ```
+
+% latex comment ?
 
@@
   "`add_q()`",           "add a column of q values to control for multiple comparisons"   
-)%>%
+) %>%
   gt(caption = "`tbl_summary()` functions to add information") %>%
@@
 
+```{r}
+data(api, package = "survey")
+
+# create weighted survey object, keep only two counties
+svy_api <- 
+  survey::svydesign(id = ~dnum, weights = ~pw, data = apiclus1, fpc = ~fpc) %>%
+  subset(cname %in% c("Alameda", "Los Angeles"))
+
+# summarize data
+tbl_svysummary_1 <-
+  svy_api %>%
+  tbl_svysummary(by = cname, 
+                 include = c(growth, target, cname),
+                 label = list(growth ~ "Growth",
+                              target ~ "Target")) %>%
+  add_p()
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_svysummary_1 %>% 
+  as_gt() %>%
+  gtsave(file = "svysummary.png")
+```
+
 ## `tbl_cross()`
+
+```{r}
+tbl_cross_1 <-
+  trial %>%
+  tbl_cross(row = stage, col = trt, percent = "cell") %>%
+  add_p(source_note = TRUE)
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_cross_1 %>% 
+  as_gt() %>%
+  gtsave(file = "cross.png")
+```
 
@@
 
+```{r}
+library(survival)
+
+tbl_survfit_1 <-
+  survfit(Surv(ttdeath, death) ~ trt, trial) %>%
+  tbl_survfit(times = c(12, 24),
+              label_header = "**{time} Month**") %>%
+  add_p()
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_cross_1 %>% 
+  as_gt() %>%
+  gtsave(file = "cross.png")
+```
+
 ## Customization
+
+```{r}
+tibble::tribble(
+  ~Function,                     ~Description,
+  "`modify_header()`",           "update column headers",   
+  "`modify_footnote()`",         "update column footnote",   
+  "`modify_spanning_header()`",  "update spanning headers",   
+  "`bold_labels()`",             "bold variable labels",  
+  "`bold_levels()`",             "bold variable levels",  
+  "`italicize_labels()`",        "italicize variable labels",  
+  "`italicize_levels()`",        "italicize variable levels",  
+  "`bold_p()`",                  "bold significant p-values"
+) %>%
+  gt() %>%
+  fmt_markdown(everything()) %>%
+  fmt_markdown(everything()) %>%
+  as_latex() %>%
+  as.character() %>%
+  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:} Functions to style and modify gtsummary tables}\\\\\n\\toprule")) %>%
+  readr::write_lines("tbl_modify.tex")
+```
 
@@
 
+```{r}
+tbl_regression_1 <-
+  glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE) %>%
+  add_global_p()
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_regression_1 %>% 
+  as_gt() %>%
+  gtsave(file = "regression.png")
+```
+
 ## `tbl_uvregression()`
 
-# In-line Reporting
+```{r}
+tbl_uvregression_1 <-
+  trial %>%
+  select(response, age, grade) %>%
+  tbl_uvregression(
+    y = response, 
+    method = glm,
+    method.args = list(family = binomial),
+    exponentiate = TRUE,
+    pvalue_fun = ~style_pvalue(., digits = 2)
+  ) %>%
+  add_nevent() %>%
+  add_global_p()
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_uvregression_1 %>% 
+  as_gt() %>%
+  gtsave(file = "uvregression.png")
+```
 
 # Merging and Stacking
+
+```{r}
+tbl1 <-
+  glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE) %>%
+  add_global_p()
+
+tbl2 <-
+  coxph(Surv(ttdeath, death) ~ age + grade, trial) %>%
+  tbl_regression(exponentiate = TRUE) %>%
+  add_global_p()
+
+tbl_merge_1 <-
+  tbl_merge(
+    tbls = list(tbl1, tbl2),
+    tab_spanner = c("**Tumor Response**", "**Time to Death**")
+  )
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_merge_1 %>% 
+  as_gt() %>%
+  gtsave(file = "merge.png")
+```
+
+# Inline Reporting
 

### 2020-12-10 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
   tbl_summary(by = trt, missing = "no") %>%
+  add_n() %>%
   add_p(test = all_continuous() ~ "t.test",
-        pvalue_fun = ~style_pvalue(., digits = 2)) %>%
-  add_n()
+        pvalue_fun = ~style_pvalue(., digits = 2))
+
 \end{example}
@@
 
+\begin{example}
+# create weighted survey object, keep only two counties
+data(api, package = "survey")
+svy_api <- 
+  survey::svydesign(id = ~dnum, weights = ~pw, data = apiclus1, fpc = ~fpc) %>%
+  subset(cname %in% c("Alameda", "Los Angeles"))
+
+# summarize data
+svy_api %>%
+  tbl_svysummary(by = cname, 
+                 include = c(growth, target, cname),
+                 label = list(growth ~ "Growth",
+                              target ~ "Target")) %>%
+  add_p()
+\end{example}
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{svysummary.png}
+  \centering
+\end{figure}
+
 \subsection{\texorpdfstring{\texttt{tbl\_cross()}}{tbl\_cross()}}
+
+\begin{example}
+trial %>%
+  tbl_cross(row = stage, col = trt, percent = "cell") %>%
+  add_p(source_note = TRUE)
+\end{example}
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{cross.png}
+  \centering
+\end{figure}
 
@@
 
+\begin{example}
+library(survival)
+
+survfit(Surv(ttdeath, death) ~ trt, trial) %>%
+  tbl_survfit(times = c(12, 24),
+              label_header = "**{time} Month**") %>%
+  add_p()
+\end{example}
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{survfit.png}
+  \centering
+\end{figure}
+
 \subsection{Customization}
+
+% adding table describing fns for customization
+\input{tbl_modify.tex}
 
@@
 
-\hypertarget{tbl_uvregression}{%
-\subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}\label{tbl_uvregression}}
+\begin{example}
+glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE) %>%
+  add_global_p()
+\end{example}
+
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{regression.png}
+  \centering
+\end{figure}
+
+\subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}
+\begin{example}
+trial %>%
+  select(response, age, grade) %>%
+  tbl_uvregression(
+    y = response, 
+    method = glm,
+    method.args = list(family = binomial),
+    exponentiate = TRUE,
+    pvalue_fun = ~style_pvalue(., digits = 2)
+  ) %>%
+  add_nevent() %>%
+  add_global_p()
+\end{example}
+
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{uvregression.png}
+  \centering
+\end{figure}
 

### 2020-12-10 rmarkdown-gtsummary.Rmd
in progress
updates

@@
     ~tibble(
-      colname = str_glue('`"{.y}"`'), 
+      colname = str_glue('`{.y}`'), 
       label = attr(.x, "label"), 
@@
   ~Argument,       ~Description,
-  "`label`",       "specify the variable labels printed in table",
-  "`type`",        "specify the variable type (e.g. continuous, categorical, etc.)",
-  "`statistic`",   "change the summary statistics presented",
-  "`digits`",      "number of digits the summary statistics will be rounded to",
-  "`missing`",     "whether to display a row with the number of missing observations",
-  "`missing_text`","text label for the missing number row",
-  "`sort`",        "change the sorting of categorical levels by frequency",
-  "`percent`",     "print column, row, or cell percentages"
+  "`label=`",       "specify the variable labels printed in table",
+  "`type=`",        "specify the variable type (e.g. continuous, categorical, etc.)",
+  "`statistic=`",   "change the summary statistics presented",
+  "`digits=`",      "number of digits the summary statistics will be rounded to",
+  "`missing=`",     "whether to display a row with the number of missing observations",
+  "`missing_text=`","text label for the missing number row",
+  "`sort=`",        "change the sorting of categorical levels by frequency",
+  "`percent=`",     "print column, row, or cell percentages"
 ) %>%
@@
 ```{r, echo=FALSE, include = FALSE}
-tbl_cross_1 %>% 
+tbl_survfit_1 %>% 
   as_gt() %>%
-  gtsave(file = "cross.png")
+  gtsave(file = "survfit.png")
 ```

### 2020-12-10 tbl_tbl_summary_args.tex
in progress
updates

@@
 \begin{longtable}{ll}
+\caption{\label{tab:}\texttt{tbl\_summary()} function arguments}\\
 \toprule
@@
 \midrule
-\texttt{label} & specify the variable labels printed in table \\ 
-\texttt{type} & specify the variable type (e.g. continuous, categorical, etc.) \\ 
-\texttt{statistic} & change the summary statistics presented \\ 
-\texttt{digits} & number of digits the summary statistics will be rounded to \\ 
-\texttt{missing} & whether to display a row with the number of missing observations \\ 
-\texttt{missing\_text} & text label for the missing number row \\ 
-\texttt{sort} & change the sorting of categorical levels by frequency \\ 
-\texttt{percent} & print column, row, or cell percentages \\ 
+\texttt{label=} & specify the variable labels printed in table \\ 
+\texttt{type=} & specify the variable type (e.g. continuous, categorical, etc.) \\ 
+\texttt{statistic=} & change the summary statistics presented \\ 
+\texttt{digits=} & number of digits the summary statistics will be rounded to \\ 
+\texttt{missing=} & whether to display a row with the number of missing observations \\ 
+\texttt{missing\_text=} & text label for the missing number row \\ 
+\texttt{sort=} & change the sorting of categorical levels by frequency \\ 
+\texttt{percent=} & print column, row, or cell percentages \\ 
 \bottomrule

### 2020-12-10 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 \title{Reproducible Summary Tables with gtsummary}
-\author{by Daniel D. Sjoberg, Karissa Whiting, Margie Hannum, Michael Curry}
+\author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry}
 
@@
 
+To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
+Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled[add ref] package. 
+These labels are displayed in the {gtsummary} tables by default.
+Using {gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.
+
 % adding table describing trial dataset
@@
 \subsection{\texorpdfstring{\texttt{tbl\_summary()}}{tbl\_summary()}}
+
+The default output from \texttt{tbl\_summary()} is meant to be publication ready.
+The \texttt{tbl\_summary()} function can take, at minimum, a data frame as the only input, and returns descriptive statistics for each column in the data frame.
+This is often the first table of clinical manuscripts and describes characteristics of the cohort under study.
+A simple example is shown below.
+Notably, by specifying the \texttt{by=} argument, you can stratify the summary table. 
+In the example below, we have split the table by the treatment a patient received.
 
@@
 \end{figure}
+
+Note the sensible defaults with this basic usage.
+
+\begin{itemize}
+\item Variable types are automatically detected so that appropriate descriptive statistics are calculated.
+\item Label attributes from the data set are automatically printed.
+\item Missing values are listed as “Unknown” in the table.
+\item Variable levels are indented and footnotes are added.
+\end{itemize}
+
+Each of the defaults may be customized using the \texttt{tbl\_summary()} function arguments.
 
@@
 \end{example}
+
+% TODO: I do not understand why this isn't being placed below the code! https://www.overleaf.com/learn/latex/Positioning_of_Figures
 \begin{figure}[h!]
@@
 
-\address{Author One\\
-  Affiliation\\
-  Address\\
-  Country\\
-  (ORCiD if desired)\\
-  \email{author1@work}}
+\address{Daniel D. Sjoberg\\
+  Memorial Sloan Kettering Cancer Center\\
+  1275 York Ave., New York, New York 10022\\
+  USA\\
+  ORCID 0000-0003-0862-2018\\
+  \email{sjobergd@mskcc.org}}
 
-\address{Author Two\\
-  Affiliation\\
-  Address\\
-  Country\\
-  (ORCiD if desired)\\
+\address{Karissa Whiting\\
+  Memorial Sloan Kettering Cancer Center\\
+  1275 York Ave., New York, New York 10022\\
+  USA\\
+  ORCID 0000-0002-4683-1868\\
   \email{author2@work}}
 
-\address{Author Three\\
-  Affiliation\\
-  Address\\
-  Country\\
-  (ORCiD if desired)\\
+\address{Michael Curry\\
+  Memorial Sloan Kettering Cancer Center\\
+  1275 York Ave., New York, New York 10022\\
+  USA\\
+  ORCID 0000-0002-0261-4044\\
   \email{author3@work}}

### 2020-12-10 rmarkdown-gtsummary.Rmd
updates

@@
   "`sort=`",        "change the sorting of categorical levels by frequency",
-  "`percent=`",     "print column, row, or cell percentages"
+  "`percent=`",     "print column, row, or cell percentages",
+  "`include=`",     "list of variables to include in summary table"
 ) %>%
@@
 tbl_survfit_1 <-
-  survfit(Surv(ttdeath, death) ~ trt, trial) %>%
+  list(survfit(Surv(ttdeath, death) ~ trt, trial),
+       survfit(Surv(ttdeath, death) ~ grade, trial)) %>%
   tbl_survfit(times = c(12, 24),

### 2020-12-11 RJtemplate_gtsummary_submission.tex

@@
 
-Note the sensible defaults with this basic usage.
+The function is highly customizable, and it is initiated with sensible default settings.
+Specifically, \texttt{tbl\_summary} detects variable types of input data and calculates descriptive statistics accordingly.
+For example, variables coded as \texttt{0/1}, \texttt{TRUE/FALSE}, and \texttt{Yes/No} are presented dichotomously.
+Additionally, \texttt{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are automatically utilized.
 
-\begin{itemize}
-\item Variable types are automatically detected so that appropriate descriptive statistics are calculated.
-\item Label attributes from the data set are automatically printed.
-\item Missing values are listed as “Unknown” in the table.
-\item Variable levels are indented and footnotes are added.
-\end{itemize}
-
-Each of the defaults may be customized using the \texttt{tbl\_summary()} function arguments.
+Default settings may be customized using the \texttt{tbl\_summary()} function arguments.
 
@@
 \input{tbl_tbl_summary_args.tex}
+
+In the example below, continuous variables are cast to \texttt{"continuous2"}, meaning the continuous summary statistics will appear on two or more rows in the table.
+The \texttt{"age"} variable's label is updated to \texttt{"Patient Age"}.
+Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument.
+The \texttt{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \texttt{missing = "no"}.
+
 \begin{example}
@@
 
+\textbf{A note about notation:}
+Throughout the gtsummary package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
+In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
+To select groups of variables, utilize the select helpers from the tidyselect[add ref] and gtsummary.
+Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
+Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse[add ref] package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
+
+The gtsummary package has several functions to add information or statistics to \texttt{tbl\_summary()} tables.
+
 % code for tbl_summary() with arguments
 \input{tbl_tbl_summary_family.tex}
+
+In the example below the number of non-missing observations is reported for each variable, as well as a p-value comparing the values between the treatment.
+The \texttt{add\_p(pvalue\_fun=)} argument accept both a proper function as well the formula shortcut notation used through the tidyverse packages.
+
 \begin{example}
@@
 \subsection{\texorpdfstring{\texttt{tbl\_svysummary()}}{tbl\_svysummary()}}
+
+The \texttt{tbl\_svysummary()} function is very similar to \texttt{tbl\_summary()} except a survey object is supplied rather than a data frame.
 
@@
 
+The \texttt{tbl\_cross()} function continues using similar syntax as the previous functions, resulting in a cross tabulation table ready for publication.
+
 \begin{example}
@@
 
+The \texttt{tbl\_survfit()} parses and tabulates n-year survival and survival percentile estimates from \texttt{survival::survfit()}.
+
 \begin{example}
@@
 
-survfit(Surv(ttdeath, death) ~ trt, trial) %>%
+list(survfit(Surv(ttdeath, death) ~ trt, trial),
+     survfit(Surv(ttdeath, death) ~ grade, trial)) %>%
   tbl_survfit(times = c(12, 24),

### 2020-12-11 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 % !TeX root = RJwrapper.tex
-\title{Reproducible Summary Tables with gtsummary}
+\title{Reproducible Summary Tables with the gtsummary Package}
 \author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry}
@@
 The \texttt{"age"} variable's label is updated to \texttt{"Patient Age"}.
-Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument.
+Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument. ADD SOMETHING ABOUT GLUE SYNTAX!
 The \texttt{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \texttt{missing = "no"}.
@@
 
-The \texttt{tbl\_survfit()} parses and tabulates n-year survival and survival percentile estimates from \texttt{survival::survfit()}.
+The \texttt{tbl\_survfit()} parses and tabulates n-year survival and survival percentile estimates from \texttt{survival::survfit()} objects.
 
@@
 
+The gtsummary package includes functions specifically made to modify and format the summary tables.
+These functions work with any table constructed with gtsummary.
+Most common uses are changing the column headers and footnotes.
+
 % adding table describing fns for customization
@@
 
+Regression modelling is one of the most common tools of medical researchers.
+The gtsummary has two functions to assist researchers and analysts to prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
+
 \subsection{\texorpdfstring{\texttt{tbl\_regression()}}{tbl\_regression()}}
+
+The \texttt{tbl\_regression()} function takes a regression model object in R and returns a formatted table of regression model results. 
+Like \texttt{tbl\_summary()}, \texttt{tbl\_regression()} creates highly customizable analytic tables with sensible defaults.
+Common regression models, such as logistic regression and Cox proportional hazards regression, are automatically identified and the tables headers are pre-filled with appropriate column headers (i.e.
+Odds Ratio and Hazard Ratio).
+
+In the example below, the logistic regression model is summarized with \texttt{tbl\regression()}.
+Note that reference row for Grade I has been added, and the variable labels have been carried through into the table.
+Using \texttt{exponentiate = TRUE} exponentiates the regression coefficients yielding the odds ratios.
+The helper function \texttt{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
 
@@
 
+The \texttt{tbl\_regression()} function leverages the huge effort behind the broom[add ref] package's \texttt{tidy()} function to perform the initial formatting of the regression object.
+Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
+The gtsummary package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.
+
 \subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}
+
+The \texttt{tbl\_uvregression()} function is a wrapper for \texttt{tbl\_regression()} that is useful when you need a series of univariate regression models.
+The user passes a data frame to \texttt{tbl\_uvregression()}, indicates what the outcome is, what regression model to run, and the function will return a beautifully formatted table of stacked univarate regression models.
+
 \begin{example}
@@
 
-\hypertarget{in-line-reporting}{%
-\section{In-line Reporting}\label{in-line-reporting}}
+\section{Inline Reporting}\label{in-line-reporting}
 
-\hypertarget{merging-and-stacking}{%
-\section{Merging and Stacking}\label{merging-and-stacking}}
+Reproducible reports are an important part of good practices.
+We often need to report the results from a table in the text of an R markdown report.
+Inline reporting has been made simple with \texttt{inline\_text()}.
+The function reports statistics from gtsummary tables inline in an R markdown document.
 
-\hypertarget{themes}{%
-\section{Themes}\label{themes}}
+Imagine you need to report the results for age from the univariate table above.
+Typically, the odds ratio, confidence interval, and p-value would be hard-coded into a report, which can lead to reproducibility issues if the data is updated and the hard-coded statistics are not amended.
+A simple call to the \texttt{inline\_text()} function will dynamically add the model results to a an Rmarkdown report.
 
-\hypertarget{print-engines}{%
-\section{Print Engines}\label{print-engines}}
+\begin{quote}
+The odds ratio for age was \texttt{\textasciigrave{}r\ inline\_text(uvreg,\ variable\ =\ age)\textasciigrave{}}.
+\end{quote}
+
+Here's how the line will appear in your report.
+
+\begin{quote}
+The odds ratio for age was 1.02 (95\% CI 1.00, 1.04; p=0.091).
+\end{quote}
+
+The default pattern to display for a regression table is \texttt{"\{estimate\} (\{conf.level*100\}\% CI \{conf.low\}, \{conf.high\}; \{p.value\})"} (again using glue syntax), and can be modified with the \texttt{inline\_text(pattern=)} argument.
+
+\section{Merging and Stacking}\label{merging-and-stacking}
+
+\section{Themes}\label{themes}
+
+\section{Print Engines}\label{print-engines}
 

### 2020-12-11 RJtemplate_gtsummary_submission.tex

@@
 
+ADD SOMETHING ABOUT as\_gt() and using the many fns available to us there.
+
 \section{Model Summaries}
@@
 
-In the example below, the logistic regression model is summarized with \texttt{tbl\regression()}.
+In the example below, the logistic regression model is summarized with \texttt{tbl\_regression()}.
 Note that reference row for Grade I has been added, and the variable labels have been carried through into the table.
@@
 
+The default styling (e.g.
+statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International.[]
+
+The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults controlling virtually every aspect of summary tables. 
+gtsummary themes were created to make these customizations easy to navigate and reuse across documents or projects. 
+Similar to ggplot themes, gtsummary themes are essentially defined sets of customization preferences that can be packaged and set for a single table or an entire document. 
+Several ready built themes are available in the package….
+
+With themes, users can control default settings for existing functions (e.g.
+always present means instead of medians in tbl_summary()), as well as many more advanced or fine-grained customizations.
+Theme elements fall into two categories: (1) elements that modify internal behavior of the functions that is not directly controllable by function arguments, (2) elements that set function argument defaults.
+The values of these theme elements must align with the functions’ accepted input for the argument.
+
+While a full glossary of customizable theme elements is available via the package documentation (http://www.danieldsjoberg.com/gtsummary/articles/themes.html)
 \section{Print Engines}\label{print-engines}

### 2020-12-11 RJtemplate_gtsummary_submission.tex

@@
 To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
-Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled[add ref] package. 
+Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package\citep{labelled}. 
 These labels are displayed in the {gtsummary} tables by default.
@@
 The default styling (e.g.
-statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International.[]
+statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International\citep{assel2019guidelines}.
 

### 2020-12-11 RJtemplate_gtsummary_submission.tex

@@
 In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
-To select groups of variables, utilize the select helpers from the tidyselect[add ref] and gtsummary.
+To select groups of variables, utilize the select helpers from the tidyselect\citep{tidyselect} and gtsummary.
 Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
-Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse[add ref] package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
+Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse\citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
 
@@
 
-The \texttt{tbl\_regression()} function leverages the huge effort behind the broom[add ref] package's \texttt{tidy()} function to perform the initial formatting of the regression object.
+The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function to perform the initial formatting of the regression object\citep{broom}.
 Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
@@
 statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International\citep{assel2019guidelines}.
+However, you'll likely submit to another journal, or your personal preferences different from the defaults.
+The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
+Themes were created to make these customizations easy to navigate and reuse across documents or projects. 
+Several ready built themes are available in the package.
+With themes, users can control default settings for existing functions (e.g.
+always present means instead of medians in \texttt{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
 
-The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults controlling virtually every aspect of summary tables. 
-gtsummary themes were created to make these customizations easy to navigate and reuse across documents or projects. 
-Similar to ggplot themes, gtsummary themes are essentially defined sets of customization preferences that can be packaged and set for a single table or an entire document. 
-Several ready built themes are available in the package….
-
-With themes, users can control default settings for existing functions (e.g.
-always present means instead of medians in tbl_summary()), as well as many more advanced or fine-grained customizations.
+many more advanced or fine-grained customizations.
 Theme elements fall into two categories: (1) elements that modify internal behavior of the functions that is not directly controllable by function arguments, (2) elements that set function argument defaults.
@@
   ORCID 0000-0002-4683-1868\\
-  \email{author2@work}}
+  \email{whitingk@mskcc.org}}
 
@@
   ORCID 0000-0002-0261-4044\\
-  \email{author3@work}}
+  \email{currym1@mskcc.org}}

### 2020-12-11 rmarkdown-gtsummary.Rmd
updates

@@
 
+To report the result for `age`, use the following commands inline.
+
+> `` `r
+inline_text(tbl_uvregression_1, variable = age)` ``
+
+Here's how the line will appear in your report.
+
+> `r inline_text(tbl_uvregression_1, variable = age)`
+
 # Themes
 
+```{r, echo=TRUE}
+theme_gtsummary_journal("nejm")
+
+tbl_nejm <-
+  glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE)
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_nejm %>% 
+  as_gt() %>%
+  gtsave(file = "nejm.png")
+```
+
 # Print Engines

### 2020-12-11 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 
-\section{Inline Reporting}\label{in-line-reporting}
+\section{Inline Reporting}
 
@@
 
-\section{Merging and Stacking}\label{merging-and-stacking}
+\section{Merging and Stacking}
 
-\section{Themes}\label{themes}
+\section{Themes}
 
@@
 
-many more advanced or fine-grained customizations.
-Theme elements fall into two categories: (1) elements that modify internal behavior of the functions that is not directly controllable by function arguments, (2) elements that set function argument defaults.
-The values of these theme elements must align with the functions’ accepted input for the argument.
+For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"}.
 
-While a full glossary of customizable theme elements is available via the package documentation (http://www.danieldsjoberg.com/gtsummary/articles/themes.html)
-\section{Print Engines}\label{print-engines}
+\begin{example}
+theme_gtsummary_journal("nejm")
+
+glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE)
+\end{example}
+
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{nejm.png}
+  \centering
+\end{figure}
+
+The themes are an evolving feature and we welcome additions of new journals and other useful themes.
+A full glossary of customizable theme elements is available in the package documentation (\url{http://www.danieldsjoberg.com/gtsummary/articles/themes.html}).
+
+\section{Print Engines}
 

### 2020-12-11 rmarkdown-gtsummary.Rmd
updates

@@
 
+```{r,echo=FALSE}
+tbl_custom <-
+  trial %>%
+  select(age, grade, trt) %>%
+  tbl_summary(by = trt, missing = "no") %>%
+  add_overall() %>%
+  add_p() %>%
+  add_stat_label() %>%
+  modify_header(label ~ "**Variable**") %>% 
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Recieved**") %>%
+  as_gt() %>%
+  gt::tab_header(
+    title = gt::md("**Table 1. Patient Characteristics**"),
+    subtitle = gt::md("_Highly Confidential_")
+  ) %>%
+  gt::tab_source_note("Data updated June 26, 2015")
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_custom %>% 
+  gtsave(file = "custom.png")
+```
+
 # Model Summaries

### 2020-12-11 RJtemplate_gtsummary_submission.tex

@@
 \end{example}
-\begin{figure}[h!]
+\begin{figure}[H]
   \includegraphics[height=4cm]{svysummary.png}
@@
 \input{tbl_modify.tex}
+
+The gtsummary package utilizes the gt package\citep{gt} to print the summary tables.
+The gt package exports approximately one hundred functions to customize and style tables.
+When you need to add additional details or styling not available within gtsummary, use the \texttt{as\_gt()} to convert the gtsummary object to gt and continue customization.
+
+\begin{example}
+
+\end{example}
 
@@
 
+SHOULD WE JUST REMOVE THIS SECTION?!
+
+\begin{figure}[h!]
+  \includegraphics[height=4.5cm]{print_engines.png}
+  \centering
+\end{figure}
+
+\section{Summary}
+
 \bibliography{RJreferences}

### 2020-12-11 RJtemplate_gtsummary_submission.tex
overleaf

@@
 \begin{example}
-
+trial %>%
+  select(age, grade, trt) %>%
+  tbl_summary(by = trt, missing = "no") %>%  
+  add_overall() %>%    
+  add_p() %>%      
+  add_stat_label() %>%     
+  modify_header(label ~ "**Variable**") %>%   
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Recieved**") %>%  
+  as_gt() %>% 
+  gt::tab_header(   
+    title = gt::md("**Table 1. Patient Characteristics**"),
+    subtitle = gt::md("_Highly Confidential_")
+  ) %>%
+  gt::tab_source_note("Data updated June 26, 2015")   
 \end{example}
-
-ADD SOMETHING ABOUT as\_gt() and using the many fns available to us there.
+\begin{figure}[h!]
+  \includegraphics[height=4cm]{custom.png}
+  \centering
+\end{figure}
 

### 2020-12-11 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 \abstract{
-An abstract of less than 150 words.
+The gtsummary package provides an elegant and flexible way to create publication-ready summary tables in R.
+The motivation behind the package stems from the work of statisticians, data scientists, and analysts, where every day, they summarize datasets and regression models in R, share results with collaborators, and eventually include them in published manuscripts or share with stakeholders.
+The gtsummary streamlines this process creating summaries of data sets, regression models, survey data, and survival data with simple code resulting in beautiful tables suitable for publication--often with a single line of code.
+The package follows a tidy framework, yielding an interface that is easy to learn that is also highly customizable.
 }
@@
 
+\citep{gtsummary}
+
 \section{Data Summaries}
@@
 To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
-Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package\citep{labelled}. 
+Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package \citep{labelled}. 
 These labels are displayed in the {gtsummary} tables by default.
@@
 \begin{figure}[h!]
-  \includegraphics[height=5cm]{summary_basic.png}
+  \includegraphics[scale=0.28]{summary_basic.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[height=5cm]{summary_plus.png}
+  \includegraphics[scale=0.28]{summary_plus.png}
   \centering
@@
 In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
-To select groups of variables, utilize the select helpers from the tidyselect\citep{tidyselect} and gtsummary.
+To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and gtsummary.
 Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
-Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse\citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
+Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
 
@@
 \begin{figure}[h!]
-  \includegraphics[height=5cm]{summary_plus_plus.png}
+  \includegraphics[scale=0.28]{summary_plus_plus.png}
   \centering
@@
 \end{example}
-\begin{figure}[H]
-  \includegraphics[height=4cm]{svysummary.png}
+\begin{figure}[h!]
+  \includegraphics[scale=0.28]{svysummary.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{cross.png}
+  \includegraphics[scale=0.28]{cross.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{survfit.png}
+  \includegraphics[scale=0.28]{survfit.png}
   \centering
@@
 
-The gtsummary package utilizes the gt package\citep{gt} to print the summary tables.
+The gtsummary package utilizes the gt package \citep{gt} to print the summary tables.
 The gt package exports approximately one hundred functions to customize and style tables.
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{custom.png}
+  \includegraphics[scale=0.28]{custom.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{regression.png}
+  \includegraphics[scale=0.28]{regression.png}
   \centering
@@
 
-The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function to perform the initial formatting of the regression object\citep{broom}.
+The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function \citep{broom} to perform the initial formatting of the regression object.
 Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{uvregression.png}
+  \includegraphics[scale=0.28]{uvregression.png}
   \centering
@@
 The default styling (e.g.
-statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International\citep{assel2019guidelines}.
+statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
 However, you'll likely submit to another journal, or your personal preferences different from the defaults.
@@
 For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"}.
+A theme was used to construct the tables shown in this manuscript to change the font to match the \emph{R Journal} font.
 
@@
 \begin{figure}[h!]
-  \includegraphics[height=4cm]{nejm.png}
+  \includegraphics[scale=0.28]{nejm.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[height=4.5cm]{print_engines.png}
+  \includegraphics[scale=0.28]{print_engines.png}
   \centering

### 2020-12-14 rmarkdown-gtsummary.Rmd
updates

--- 2020-12-11 rmarkdown-gtsummary.Rmd
@@
 
+```{r,echo=FALSE}
+tbl_custom <-
+  trial %>%
+  select(age, grade, trt) %>%
+  tbl_summary(by = trt, missing = "no") %>%
+  add_overall() %>%
+  add_p() %>%
+  add_stat_label() %>%
+  modify_header(label ~ "**Variable**") %>% 
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Received**") %>%
+  as_gt() %>%
+  gt::tab_header(
+    title = gt::md("**Table 1. Patient Characteristics**"),
+    subtitle = gt::md("_Highly Confidential_")
+  ) %>%
+  gt::tab_source_note("Data updated June 26, 2015")
+```
+
+```{r, echo=FALSE, include = FALSE}
+tbl_custom %>% 
+  gtsave(file = "custom.png")
+```
+
 # Model Summaries

### 2020-12-14 RJtemplate_gtsummary_submission.tex
overleaf
Updates from Overleaf

@@
 % !TeX root = RJwrapper.tex
-\title{Reproducible Summary Tables with the gtsummary Package}
+\title{Reproducible Summary Tables with the e \pkg{gtsummary} Package}
 \author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry}
@@
 \abstract{
-The gtsummary package provides an elegant and flexible way to create publication-ready summary tables in R.
+The \pkg{gtsummary} package provides an elegant and flexible way to create publication-ready summary tables in R.
 The motivation behind the package stems from the work of statisticians, data scientists, and analysts, where every day, they summarize datasets and regression models in R, share results with collaborators, and eventually include them in published manuscripts or share with stakeholders.
-The gtsummary streamlines this process creating summaries of data sets, regression models, survey data, and survival data with simple code resulting in beautiful tables suitable for publication--often with a single line of code.
+The \pkg{gtsummary} streamlines this process creating summaries of data sets, regression models, survey data, and survival data with simple code resulting in beautiful tables suitable for publication--often with a single line of code.
 The package follows a tidy framework, yielding an interface that is easy to learn that is also highly customizable.
@@
 \section{Introduction}
+A common need in research and business analytics is summarizing data and statistical models.
+Table summaries identify patterns in the data that help analysts discover insights and drive action.
+While there are options to create summary tables in R, few, if any packages, have the ability to make tables that are easy to: export, interpret, reproduce, and create.
 
-\citep{gtsummary}
+The \CRANpkg{gtsummary} \citep{gtsummary} package provides an elegant and flexible way to create publication-ready analytical and summary tables in R.
+This package works with existing packages and frameworks to close the loop between reproducible RMarkdown R code and polished, organized statistical reports.
+Specifically, gtsummary allows the users to fully customize the format of tables in R so users do not have to modify any tables after an RMarkdown has rendered or a table has been exported.
+Getting rid of the need to modify tables after the table has been created, reduces an error-prone workflow and increases the chances that your analysis is reproducible.
+
+Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize the most common types of regression models, and report statistics in line in R markdown reports.
+Identifying building blocks of tables presented in medical literature (that also are common among all fields), \pkg{gtsummary} seeks to create these tables and their formats so there is no need for further end-user modifications.
+
+Additionally, \pkg{gtsummary} leverages other data cleaning and analysis packages to create a complete analysis and reporting framework.
+For example, users can take advantage of existing broom tidiers to clean regression results for \texttt{tbl\_regression()} and use \pkg{gt}, \pkg{kable}, \pkg{flextable} or others to print \pkg{gtsummary} tables to various output formats (e.g.
+HTML, PDF, Word, RTF).
+Furthermore, all \pkg{gtsummary} functions are created to work within a “tidy” framework where a data frame is the first argument and functions can be used with pipes.
+Unquoted variables are used and tidyselect functions can be used to select and manipulate groups of variables within the framework.
+
+While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
+For example, \pkg{finalfit} only creates regressions outputs and you have to manually specify a printing engine for the table to render.
+Furthermore, \pkg{tableone} reports on general descriptive statistics, it is fundamentally lacking in its ability to produce tables for regression modeling strategies in R.
+
+\pkg{gtsummary} comes equipped common descriptive summaries for several data types allowing the user to easily aggregate data.
+At first, \pkg{gtsummary} will identify data as continuous or discrete and apply summary methods to each variable. 
+However, any default in identifying data types can be manually overridden. 
+Additionally, if any descriptive statistical method is not inherently in \pkg{gtsummary}, the user can supply their own function to summarise the data. 
+Along with common descriptive summaries, \pkg{gtsummary} has the most common methods to summarise statistical models, survey data, and even perform cross tables for comparing discrete variables. 
+After data summarization, \pkg{gtsummary} allows users to combine tables, either side by side (with \texttt{tbl\_merge()}) , or on top of each other (with \texttt{tbl\_stack()}). 
+This data merging allows users to easily synthesize and compare output from several tables and share a lot of information in a compact format.
 
@@
 
-To show use of gtsummary functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
+To show use of \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
 Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package \citep{labelled}. 
-These labels are displayed in the {gtsummary} tables by default.
-Using {gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.
+These labels are displayed in the \pkg{gtsummary} tables by default.
+Using \pkg{gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.
 
@@
 \textbf{A note about notation:}
-Throughout the gtsummary package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
+Throughout the \pkg{gtsummary} package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
 In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
-To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and gtsummary.
+To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and \pkg{gtsummary}.
 Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
-Similarly, users may utilize \texttt{all\_categorical()} (from gtsummary), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
+Similarly, users may utilize \texttt{all\_categorical()} (from \pkg{gtsummary}), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
 
-The gtsummary package has several functions to add information or statistics to \texttt{tbl\_summary()} tables.
+The \pkg{gtsummary} package has several functions to add information or statistics to \texttt{tbl\_summary()} tables.
 
@@
 
-The gtsummary package includes functions specifically made to modify and format the summary tables.
-These functions work with any table constructed with gtsummary.
+The \pkg{gtsummary} package includes functions specifically made to modify and format the summary tables.
+These functions work with any table constructed with \pkg{gtsummary}.
 Most common uses are changing the column headers and footnotes.
@@
 
-The gtsummary package utilizes the gt package \citep{gt} to print the summary tables.
+The \pkg{gtsummary} package utilizes the gt package \citep{gt} to print the summary tables.
 The gt package exports approximately one hundred functions to customize and style tables.
-When you need to add additional details or styling not available within gtsummary, use the \texttt{as\_gt()} to convert the gtsummary object to gt and continue customization.
+When you need to add additional details or styling not available within \pkg{gtsummary}, use the \texttt{as\_gt()} to convert the \pkg{gtsummary} object to gt and continue customization.
 
@@
   modify_header(label ~ "**Variable**") %>%   
-  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Recieved**") %>%  
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Received**") %>%  
   as_gt() %>% 
@@
 Regression modelling is one of the most common tools of medical researchers.
-The gtsummary has two functions to assist researchers and analysts to prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
+The \pkg{gtsummary} has two functions to assist researchers and analysts to prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
 
@@
 Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
-The gtsummary package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.
+The \pkg{gtsummary} package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.
 
@@
 Inline reporting has been made simple with \texttt{inline\_text()}.
-The function reports statistics from gtsummary tables inline in an R markdown document.
+The function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
 
@@
 However, you'll likely submit to another journal, or your personal preferences different from the defaults.
-The gtsummary package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
+The \pkg{gtsummary} package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
 Themes were created to make these customizations easy to navigate and reuse across documents or projects. 
@@
 
+The \pkg{gtsummary} package 
 SHOULD WE JUST REMOVE THIS SECTION?!
@@
   USA\\
-  ORCID 0000-0003-0862-2018\\
+  ORCiD: 0000-0003-0862-2018\\
   \email{sjobergd@mskcc.org}}
@@
   USA\\
-  ORCID 0000-0002-4683-1868\\
+  ORCiD: 0000-0002-4683-1868\\
   \email{whitingk@mskcc.org}}
@@
   USA\\
-  ORCID 0000-0002-0261-4044\\
+  ORCiD: 0000-0002-0261-4044\\
   \email{currym1@mskcc.org}}

### 2020-12-14 RJtemplate_gtsummary_submission.tex
overleaf

@@
 % !TeX root = RJwrapper.tex
-\title{Reproducible Summary Tables with the e \pkg{gtsummary} Package}
+\title{Reproducible Summary Tables with the \pkg{gtsummary} Package}
 \author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry}
@@
 
-While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
+While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique https://www.overleaf.com/project/5fd126ae12e0e68e74c14832in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
 For example, \pkg{finalfit} only creates regressions outputs and you have to manually specify a printing engine for the table to render.
@@
 
+The \pkg{gtsummary} tables shown above are often ready for publication as they are.
+It is, however, common that higher quality tables can be constructed by merging or stacking simple tables.
+For example in cancer research, we often report on models about both a tumor's response to a treatment, and if that treatment reduces the risk of death.
+These two models are often reported side-by-side, and it's simple to construct this table using  the \texttt{tbl\_merge()} function.
+Simply, build a table for each regression model using \texttt{tbl\_regression()}, then merge the to tables with \texttt{tbl\_merge()}: any number of \pkg{gtsummary} tables can be merged with this function.
+
+\begin{example}
+tbl1 <- 
+  glm(response ~ age + grade, trial, family = binomial) %>%
+  tbl_regression(exponentiate = TRUE)
+
+tbl2 <-
+  coxph(Surv(ttdeath, death) ~ age + grade, trial) %>%
+  tbl_regression(exponentiate = TRUE)
+
+tbl_merge_1 <-
+  tbl_merge(
+    tbls = list(tbl1, tbl2),
+    tab_spanner = c("**Tumor Response**", "**Time to Death**")
+  )
+\end{example}
+
+\begin{figure}[h!]
+  \includegraphics[scale=0.28]{merge.png}
+  \centering
+\end{figure}
+
+Similarly, any number of \pkg{gtsummary} tables may be stacked on top of one another using the \texttt{tbl\_stack()} function.
+
 \section{Themes}
@@
 statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
-However, you'll likely submit to another journal, or your personal preferences different from the defaults.
+However, you'll likely submit to another journal, or your personal preferences differ from the defaults.
 The \pkg{gtsummary} package is unique from other table building packages in the ability to set fine-grained customization defaults with themes. 
 Themes were created to make these customizations easy to navigate and reuse across documents or projects. 
-Several ready built themes are available in the package.
 With themes, users can control default settings for existing functions (e.g.
 always present means instead of medians in \texttt{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
+Several themes are available in the package that follow various journals' reporting guidelines, reduce cell padding, and recently language themes were introduced and the output of \pkg{gtsummary} tables can be translated into more than 10 languages.
 
-For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"}.
-A theme was used to construct the tables shown in this manuscript to change the font to match the \emph{R Journal} font.
+For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"} (ampong other modifications).
 
@@
 
-The themes are an evolving feature and we welcome additions of new journals and other useful themes.
-A full glossary of customizable theme elements is available in the package documentation (\url{http://www.danieldsjoberg.com/gtsummary/articles/themes.html}).
+A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font. 
+Themes are an evolving feature and we welcome additions of new journals' reporting guidelines and other themes useful to users.
+A full glossary of customizable theme elements is available in the package's themes vignette (\url{http://www.danieldsjoberg.com/gtsummary/articles/themes.html}).
 

### 2020-12-16 RJtemplate_gtsummary_submission.tex
overleaf

@@
 \maketitle
+
+% TO DO
+% - change all code refs from \texttt{} to \code{}
+% - update all references to packages to use \CRANpkg{} and \pkg{}
 
@@
 Additionally, \pkg{gtsummary} leverages other data cleaning and analysis packages to create a complete analysis and reporting framework.
-For example, users can take advantage of existing broom tidiers to clean regression results for \texttt{tbl\_regression()} and use \pkg{gt}, \pkg{kable}, \pkg{flextable} or others to print \pkg{gtsummary} tables to various output formats (e.g.
-HTML, PDF, Word, RTF).
+For example, users can take advantage of existing \CRANpkg{broom} tidiers to clean regression results for \texttt{tbl\_regression()} and use \CRANpkg{gt}, \CRANpkg{kable}, \CRANpkg{flextable} or others to print \pkg{gtsummary} tables to various output formats (e.g.
+HTML, PDF, Word, or RTF).
 Furthermore, all \pkg{gtsummary} functions are created to work within a “tidy” framework where a data frame is the first argument and functions can be used with pipes.
@@
 
-While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique https://www.overleaf.com/project/5fd126ae12e0e68e74c14832in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
-For example, \pkg{finalfit} only creates regressions outputs and you have to manually specify a printing engine for the table to render.
-Furthermore, \pkg{tableone} reports on general descriptive statistics, it is fundamentally lacking in its ability to produce tables for regression modeling strategies in R.
+While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
+For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
+Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, it is fundamentally lacking in its ability to produce tables for regression modeling strategies in R.
 
@@
 The \texttt{"age"} variable's label is updated to \texttt{"Patient Age"}.
-Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument. ADD SOMETHING ABOUT GLUE SYNTAX!
+Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument. 
+\pkg{gtsummary} uses \CRANpkg{glue} to help interpret string literals. Glue embeds R expressions (functions or objects) in curly braces which are then evaluated and inserted into the argument string. 
+ADD SOMETHING ABOUT GLUE SYNTAX!
 The \texttt{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \texttt{missing = "no"}.
@@
 
+Keep this section (or add it somewhere else).  We'll say that the other print engines are available. We won't even mention the defaults since they are will be changing in the future.
+
 \begin{figure}[h!]

### 2020-12-16 rmarkdown-gtsummary.Rmd
updates

@@
     by = trt,
-    type = all_continuous() ~ "continuous2",
+    type = age ~ "continuous2",
     label = age ~ "Patient Age",
-    statistic = list(all_continuous() ~ c("{N_nonmiss}", 
-                                          "{mean} ({sd})", 
-                                          "{median} ({p25}, {p75})", 
-                                          "{min}, {max}"),
-                     all_categorical() ~ "{n} / {N} ({p}%)"),
-    digits = all_categorical() ~ c(0, 0, 1),
+    statistic = list(age ~ c("{N_nonmiss}", "{mean} ({sd})"),
+                     c(grade, response) ~ "{n} / {N} ({p}%)"),
+    digits = c(grade, response) ~ c(0, 0, 1),
     missing = "no"

### 2020-12-16 rmarkdown-gtsummary.Rmd
updates

@@
 ```{r}
-data(api, package = "survey")
+# convert trial data frame to survey object
+svy_trial <- survey::svydesign(data = trial, ids = ~ 1, weights = ~ 1)
 
-# create weighted survey object, keep only two counties
-svy_api <- 
-  survey::svydesign(id = ~dnum, weights = ~pw, data = apiclus1, fpc = ~fpc) %>%
-  subset(cname %in% c("Alameda", "Los Angeles"))
-
-# summarize data
 tbl_svysummary_1 <-
-  svy_api %>%
-  tbl_svysummary(by = cname, 
-                 include = c(growth, target, cname),
-                 label = list(growth ~ "Growth",
-                              target ~ "Target")) %>%
+  svy_trial %>%
+  tbl_svysummary(by = trt, include = c(trt, age, grade)) %>%
   add_p()
@@
   trial %>%
-  select(age, grade, trt) %>%
+  select(marker, grade, trt) %>%
   tbl_summary(by = trt, missing = "no") %>%
@@
   add_p() %>%
+  add_q() %>%
   add_stat_label() %>%
+  bold_p(t = 0.10) %>%
+  bold_labels() %>%
   modify_header(label ~ "**Variable**") %>% 

### 2020-12-16 RJtemplate_gtsummary_submission.tex

@@
 The \pkg{gtsummary} package provides an elegant and flexible way to create publication-ready summary tables in R.
-The motivation behind the package stems from the work of statisticians, data scientists, and analysts, where every day, they summarize datasets and regression models in R, share results with collaborators, and eventually include them in published manuscripts or share with stakeholders.
-The \pkg{gtsummary} streamlines this process creating summaries of data sets, regression models, survey data, and survival data with simple code resulting in beautiful tables suitable for publication--often with a single line of code.
-The package follows a tidy framework, yielding an interface that is easy to learn that is also highly customizable.
+A critical part of the work of statisticians, data scientists, and analysts is summarizing data sets and regression models in R and publishing or sharing polished summary tables to describe analysis results.
+The \pkg{gtsummary} package was created to streamline these everyday analysis tasks by allowing users to easily create reproducible summaries of data sets, regression models, survey data, and survival data with a simple interface and very little code.
+The package follows a tidy framework, making it easy to integrate in standard data workflows, and offers many table customization features through function arguments, helper functions and custom themes.
 }

### 2020-12-18 rmarkdown-gtsummary.Rmd
updates

@@
   "`add_n()`",           "add a column with N (or N missing) for each variable",
-  "`add_difference()`",  "add column for difference between two group, confidemce interval, and p-value",
+  "`add_difference()`",  "add column for difference between two group, confidence interval, and p-value",
   "`add_stat_label()`",  "add label for the summary statistics shown in each row",   

### 2020-12-18 tbl_tbl_summary_family.tex
in progress
updates

@@
 \begin{longtable}{ll}
-\caption{\label{tab:}\texttt{tbl_summary()} functions to add information}\\
+\caption{\label{tab:}\texttt{tbl\_summary()} functions to add information}\\
 \toprule
@@
 \texttt{add\_n()} & add a column with N (or N missing) for each variable \\ 
-\texttt{add\_difference()} & add column for difference between two group, confidemce interval, and p-value \\ 
+\texttt{add\_difference()} & add column for difference between two group, confidence interval, and p-value \\ 
 \texttt{add\_stat\_label()} & add label for the summary statistics shown in each row \\ 

### 2020-12-18 RJtemplate_gtsummary_submission.tex

@@
 To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and \pkg{gtsummary}.
-Above, \texttt{all\_continuous()} was utilized to change the summary statistics for all continuous variables. 
+The \texttt{all\_continuous()} selector is a convenient way to select all continuous variables.
+In the example above it could have been used to change the summary statistics for all continuous variables---\code{all\_continuous() ~ c("\{N\_nonmiss\}", "\{mean\} (\{sd\})")}. 
 Similarly, users may utilize \texttt{all\_categorical()} (from \pkg{gtsummary}), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
@@
 \begin{example}
-# create weighted survey object, keep only two counties
-data(api, package = "survey")
-svy_api <- 
-  survey::svydesign(id = ~dnum, weights = ~pw, data = apiclus1, fpc = ~fpc) %>%
-  subset(cname %in% c("Alameda", "Los Angeles"))
+# convert trial data frame to survey object
+svy_trial <- survey::svydesign(data = trial, ids = ~ 1, weights = ~ 1)
 
-# summarize data
-svy_api %>%
-  tbl_svysummary(by = cname, 
-                 include = c(growth, target, cname),
-                 label = list(growth ~ "Growth",
-                              target ~ "Target")) %>%
+tbl_svysummary_1 <-
+  svy_trial %>%
+  tbl_svysummary(by = trt, include = c(trt, age, grade)) %>%
   add_p()
@@
 
+The example below includes customization using both gtsummary and gt functions.
+The gtsummary functions are utilized to bold p-values below a threshold, bold the variable labels,  update the column header, and add a spanning header
+
 \begin{example}

### 2020-12-18 RJtemplate_gtsummary_submission.tex
Updates from Overleaf

@@
 % - update all references to packages to use \CRANpkg{} and \pkg{}
+% - scale the images approriately (they are too small now)
 
@@
 \section{Introduction}
-A common need in research and business analytics is summarizing data and statistical models.
-Table summaries identify patterns in the data that help analysts discover insights and drive action.
-While there are options to create summary tables in R, few, if any packages, have the ability to make tables that are easy to: export, interpret, reproduce, and create.
+
+Table summaries are a fundamental tool in an analyst's toolbox, that help them understand and communicate patterns in their data.
+While several packages exist to create summary tables in R, few, if any, have the ability to make tables that are easy to: create, export, interpret, publish and reproduce.
 
@@
 Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize the most common types of regression models, and report statistics in line in R markdown reports.
-Identifying building blocks of tables presented in medical literature (that also are common among all fields), \pkg{gtsummary} seeks to create these tables and their formats so there is no need for further end-user modifications.
+Identifying building blocks of tables presented in medical literature (that also are common among all fields), \pkg{gtsummary} seeks to create these fully-formatted tables so there is no need for further end-user modifications.
 
@@
 HTML, PDF, Word, or RTF).
-Furthermore, all \pkg{gtsummary} functions are created to work within a “tidy” framework where a data frame is the first argument and functions can be used with pipes.
+Furthermore, all \pkg{gtsummary} functions are created to work within a “tidy” framework where functions can be used with pipes.
 Unquoted variables are used and tidyselect functions can be used to select and manipulate groups of variables within the framework.
@@
 For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
-Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, it is fundamentally lacking in its ability to produce tables for regression modeling strategies in R.
+Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, but it is lacking in its ability to produce tables for regression modeling strategies in R.
 
@@
 
-In the example below, continuous variables are cast to \texttt{"continuous2"}, meaning the continuous summary statistics will appear on two or more rows in the table.
+By default, tables display one row of statistics per variable.
+This can be customized, and in the example below, the age variable is cast to \texttt{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table.
+This allows the numner of non-missing and the mean to be displayed on separate lines.
 The \texttt{"age"} variable's label is updated to \texttt{"Patient Age"}.
@@
 
-The \pkg{gtsummary} package has several functions to add information or statistics to \texttt{tbl\_summary()} tables.
+In addition to basic summary statistics, the \pkg{gtsummary} package has several functions to add additional information or statistics to \texttt{tbl\_summary()} tables.
 
@@
 In the example below the number of non-missing observations is reported for each variable, as well as a p-value comparing the values between the treatment.
-The \texttt{add\_p(pvalue\_fun=)} argument accept both a proper function as well the formula shortcut notation used through the tidyverse packages.
+Default statistical tests are chosen based on data type, but test type can be customized within the \texttt{add\_p()} function.
+P-value formatting can be adjusted using the \texttt{pvalue\_fun=} argument which accepts both a proper function, as well the formula shortcut notation used throughout tidyverse packages.
 
@@
 
-The \texttt{tbl\_cross()} function continues using similar syntax as the previous functions, resulting in a cross tabulation table ready for publication.
+The \texttt{tbl\_cross()} function is a wrapper for \texttt{tbl\_summary()} that creates a simple, publication-ready cross tabulation table.
+The syntax is consistent with\texttt{tbl\_summary()} and \texttt{tbl\_svyummary()}
 
@@
 
-The \texttt{tbl\_survfit()} parses and tabulates n-year survival and survival percentile estimates from \texttt{survival::survfit()} objects.
+The \texttt{tbl\_survfit()} function parses and tabulates \texttt{survival::survfit()} objects presenting survival percentile estimates and time point-specific survival probability estimates.
 
@@
 Regression modelling is one of the most common tools of medical researchers.
-The \pkg{gtsummary} has two functions to assist researchers and analysts to prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
+The \pkg{gtsummary} has two functions to assist researchers and analysts prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
 
@@
 In the example below, the logistic regression model is summarized with \texttt{tbl\_regression()}.
-Note that reference row for Grade I has been added, and the variable labels have been carried through into the table.
-Using \texttt{exponentiate = TRUE} exponentiates the regression coefficients yielding the odds ratios.
+Note that a reference row for Grade I has been added, and the variable labels have been carried through into the table.
+Using \texttt{exponentiate = TRUE}  weexponentiate the regression coefficients, yielding the odds ratios.
 The helper function \texttt{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
@@
 The \texttt{tbl\_uvregression()} function is a wrapper for \texttt{tbl\_regression()} that is useful when you need a series of univariate regression models.
-The user passes a data frame to \texttt{tbl\_uvregression()}, indicates what the outcome is, what regression model to run, and the function will return a beautifully formatted table of stacked univarate regression models.
+The user passes a data frame to \texttt{tbl\_uvregression()}, indicates what the outcome is, what regression model to run, and the function will return a formatted table of stacked univarate regression models.
 
@@
 
-Reproducible reports are an important part of good practices.
+Reproducible reports are an important part of good analysis practices.
 We often need to report the results from a table in the text of an R markdown report.
-Inline reporting has been made simple with \texttt{inline\_text()}.
-The function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
+The\texttt{inline\_text()}  function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
 
@@
 
-The \pkg{gtsummary} tables shown above are often ready for publication as they are.
-It is, however, common that higher quality tables can be constructed by merging or stacking simple tables.
-For example in cancer research, we often report on models about both a tumor's response to a treatment, and if that treatment reduces the risk of death.
+The \pkg{gtsummary} tables shown above are often ready for publication as they are, however it is common that higher quality tables can be constructed by merging or stacking simple tables.
+For example in cancer research, we often report on models about both a tumor's response to a treatment, and a treatment efficacy in reducing the risk of death.
 These two models are often reported side-by-side, and it's simple to construct this table using  the \texttt{tbl\_merge()} function.
-Simply, build a table for each regression model using \texttt{tbl\_regression()}, then merge the to tables with \texttt{tbl\_merge()}: any number of \pkg{gtsummary} tables can be merged with this function.
+First, build a table for each regression model using \texttt{tbl\_regression()}, then merge the two tables with \texttt{tbl\_merge()}.
+Any number of \pkg{gtsummary} tables can be merged with this function.
 
@@
 
-The \pkg{gtsummary} package 
-SHOULD WE JUST REMOVE THIS SECTION?!
-
-Keep this section (or add it somewhere else).  We'll say that the other print engines are available. We won't even mention the defaults since they are will be changing in the future.
-
-\begin{figure}[h!]
-  \includegraphics[scale=0.28]{print_engines.png}
-  \centering
-\end{figure}
+Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \texttt{knitr::kable()}, \pkg{gtsummary}, \pkg{huxtable} and more.
+While \pkg{gt} is used as default for most outputs, you can easily convert to using print engine of choice using several conversion helper functions provided in the package (e.g.
+\texttt{as\_kable()}).
+The package is designed to interact with these print engines behind the scenes to reduce burden on users and you generally only need to be aware of them if you want to add advanced customizations (eg.
+example X with \texttt{tab\_header}).
 

### 2020-12-18 rmarkdown-gtsummary.Rmd
updates

@@
   glm(response ~ age + grade, trial, family = binomial) %>%
-  tbl_regression(exponentiate = TRUE) %>%
-  add_global_p()
+  tbl_regression(exponentiate = TRUE)
 
@@
   coxph(Surv(ttdeath, death) ~ age + grade, trial) %>%
-  tbl_regression(exponentiate = TRUE) %>%
-  add_global_p()
+  tbl_regression(exponentiate = TRUE)
 

### 2020-12-18 rmarkdown-gtsummary.Rmd
updates

@@
 
-theme_gtsummary_compact()
-set_gtsummary_theme(list(
-  "as_gt-lst:addl_cmds" = list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino")))
-))
+# theme_gtsummary_compact()
+# set_gtsummary_theme(list(
+#   "as_gt-lst:addl_cmds" = list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino")))
+# ))
+
+my_r_journal_theme <- theme_gtsummary_compact(set_theme = FALSE)
+my_r_journal_theme$`as_gt-lst:addl_cmds` <-
+  my_r_journal_theme$`as_gt-lst:addl_cmds` %>% 
+  c(list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino"))))
+
+set_gtsummary_theme(my_r_journal_theme)
 ```

### 2020-12-18 RJtemplate_gtsummary_submission.tex

@@
 % TO DO
-% - change all code refs from \texttt{} to \code{}
+% - change all code refs from \code{} to \code{}
 % - update all references to packages to use \CRANpkg{} and \pkg{}
-% - scale the images approriately (they are too small now)
+% - scale the images appropriately (they are too small now)
 
@@
 The \pkg{gtsummary} package provides an elegant and flexible way to create publication-ready summary tables in R.
-A critical part of the work of statisticians, data scientists, and analysts is summarizing data sets and regression models in R and publishing or sharing polished summary tables to describe analysis results.
+A critical part of the work of statisticians, data scientists, and analysts is summarizing data sets and regression models in R and publishing or sharing polished summary tables.
 The \pkg{gtsummary} package was created to streamline these everyday analysis tasks by allowing users to easily create reproducible summaries of data sets, regression models, survey data, and survival data with a simple interface and very little code.
-The package follows a tidy framework, making it easy to integrate in standard data workflows, and offers many table customization features through function arguments, helper functions and custom themes.
+The package follows a tidy framework, making it easy to integrate in standard data workflows, and offers many table customization features through function arguments, helper functions, and custom themes.
 }
@@
 
-Table summaries are a fundamental tool in an analyst's toolbox, that help them understand and communicate patterns in their data.
-While several packages exist to create summary tables in R, few, if any, have the ability to make tables that are easy to: create, export, interpret, publish and reproduce.
+Table summaries are a fundamental tool in an analyst's toolbox that help us understand and communicate patterns in our data,
+The ability to easily create and export polished and reproducible tables is essential.
+The \CRANpkg{gtsummary} \citep{gtsummary} package provides an elegant and flexible framework to create publication-ready analytical and summary tables in R.
+This package works to gap the loop between a reproducible RMarkdown report and the final report.
+Specifically, \pkg{gtsummary} allows the us to fully customize and format summary tables with code, eliminating the need to modify any tables by hand after the table has been exported.
+Removing the need to modify tables after the table has been created, eliminates an error-prone step in our workflow and increases the reproducibility of our analyses and reports.
 
-The \CRANpkg{gtsummary} \citep{gtsummary} package provides an elegant and flexible way to create publication-ready analytical and summary tables in R.
-This package works with existing packages and frameworks to close the loop between reproducible RMarkdown R code and polished, organized statistical reports.
-Specifically, gtsummary allows the users to fully customize the format of tables in R so users do not have to modify any tables after an RMarkdown has rendered or a table has been exported.
-Getting rid of the need to modify tables after the table has been created, reduces an error-prone workflow and increases the chances that your analysis is reproducible.
+Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize regression models, and report statistics inline in RMarkdown reports.
+After identifying these basic structures of most tables presented in medical literature (and other fields), we created \pkg{gtsummary} to ease creation of fully-formatted, ready to publish tables.
 
-Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize the most common types of regression models, and report statistics in line in R markdown reports.
-Identifying building blocks of tables presented in medical literature (that also are common among all fields), \pkg{gtsummary} seeks to create these fully-formatted tables so there is no need for further end-user modifications.
+Additionally, \pkg{gtsummary} leverages other analysis and tidying R packages to create a complete analysis and reporting framework.
+For example, we take advantage of the existing \CRANpkg{broom} tidiers to prepare regression results for \code{tbl\_regression()} and use \CRANpkg{gt} \citep{gt} to print \pkg{gtsummary} tables to various output formats (e.g.
+HTML, PDF, Word, or RTF).
+Furthermore, \pkg{gtsummary} functions are designed to work within a "tidy" framework, utilizing the \CRANpkg{magrittr} pipe operator and \CRANpkg{tidyselect} \citep{tidyselect} functions used throughout the \CRANpkg{tidyverse} \citep{tidyverse}.
 
-Additionally, \pkg{gtsummary} leverages other data cleaning and analysis packages to create a complete analysis and reporting framework.
-For example, users can take advantage of existing \CRANpkg{broom} tidiers to clean regression results for \texttt{tbl\_regression()} and use \CRANpkg{gt}, \CRANpkg{kable}, \CRANpkg{flextable} or others to print \pkg{gtsummary} tables to various output formats (e.g.
-HTML, PDF, Word, or RTF).
-Furthermore, all \pkg{gtsummary} functions are created to work within a “tidy” framework where functions can be used with pipes.
-Unquoted variables are used and tidyselect functions can be used to select and manipulate groups of variables within the framework.
+While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables,offers diverse features to customize content of tables to a high degree, and includes the ability to report statistics from the tables inline in an RMarkdown report.
+%For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
+%Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, but it is lacking in its ability to produce tables for regression modeling strategies in R.
 
-While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for all types of statistical tables and offers diverse features to customize content and format of tables to a high degree.
-For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
-Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, but it is lacking in its ability to produce tables for regression modeling strategies in R.
-
-\pkg{gtsummary} comes equipped common descriptive summaries for several data types allowing the user to easily aggregate data.
-At first, \pkg{gtsummary} will identify data as continuous or discrete and apply summary methods to each variable. 
-However, any default in identifying data types can be manually overridden. 
-Additionally, if any descriptive statistical method is not inherently in \pkg{gtsummary}, the user can supply their own function to summarise the data. 
-Along with common descriptive summaries, \pkg{gtsummary} has the most common methods to summarise statistical models, survey data, and even perform cross tables for comparing discrete variables. 
-After data summarization, \pkg{gtsummary} allows users to combine tables, either side by side (with \texttt{tbl\_merge()}) , or on top of each other (with \texttt{tbl\_stack()}). 
-This data merging allows users to easily synthesize and compare output from several tables and share a lot of information in a compact format.
+Along with descriptive summaries, \pkg{gtsummary} summarises statistical models, survey data, and build cross tabulations. 
+After data are summarized in a table, \pkg{gtsummary} allows users to combine tables, either side-by-side (with \code{tbl\_merge()}) , or on top of each other (with \code{tbl\_stack()}). 
+This table merging and stacking allows us to easily synthesize and compare output from several tables and share information in a compact format.
 
@@
 
-To show use of \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcome of tumor response to the treatment.
-Each variable in the data frame has been assigned an attribute label (i.e. \texttt{attr(trial\$trt, "label") == "Chemotherapy Treatment"}) with the labelled package \citep{labelled}. 
-These labels are displayed in the \pkg{gtsummary} tables by default.
-Using \pkg{gtsummary} on a data frame without labels will print variable names in place of variable labels with the option to update the text displayed.
+To showcase \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcomes of tumor response and death.
+Each variable in the data frame has been assigned an attribute label with the \pkg{labelled} package \citep{labelled}, e.g.
+\code{trial \%>\% set\_variable\_labels(age = "Age")}, that will be shown in the summary tables.
+These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned the variable name would have been shown.
 
@@
 
-\subsection{\texorpdfstring{\texttt{tbl\_summary()}}{tbl\_summary()}}
+\subsection{\texorpdfstring{\code{tbl\_summary()}}{tbl\_summary()}}
 
-The default output from \texttt{tbl\_summary()} is meant to be publication ready.
-The \texttt{tbl\_summary()} function can take, at minimum, a data frame as the only input, and returns descriptive statistics for each column in the data frame.
-This is often the first table of clinical manuscripts and describes characteristics of the cohort under study.
+The default output from \code{tbl\_summary()} is meant to be publication ready.
+The \code{tbl\_summary()} function can take, at minimum, a data frame as the only input, and returns descriptive statistics for each column in the data frame.
+This is often the first table of clinical manuscripts and describes characteristics of the study cohort.
 A simple example is shown below.
-Notably, by specifying the \texttt{by=} argument, you can stratify the summary table. 
+Notably, by specifying the \code{by=} argument, you can stratify the summary table. 
 In the example below, we have split the table by the treatment a patient received.
@@
 % code for basic tbl_summary()
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{summary_basic.png}
+  \includegraphics[scale=0.35]{summary_basic.png}
   \centering
@@
 The function is highly customizable, and it is initiated with sensible default settings.
-Specifically, \texttt{tbl\_summary} detects variable types of input data and calculates descriptive statistics accordingly.
-For example, variables coded as \texttt{0/1}, \texttt{TRUE/FALSE}, and \texttt{Yes/No} are presented dichotomously.
-Additionally, \texttt{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are automatically utilized.
+Specifically, \code{tbl\_summary} detects variable types of input data and calculates descriptive statistics accordingly.
+For example, variables coded as \code{0/1}, \code{TRUE/FALSE}, and \code{Yes/No} are presented dichotomously.
+Additionally, \code{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are automatically utilized.
 
-Default settings may be customized using the \texttt{tbl\_summary()} function arguments.
+Default settings may be customized using the \code{tbl\_summary()} function arguments.
 
@@
 
-By default, tables display one row of statistics per variable.
-This can be customized, and in the example below, the age variable is cast to \texttt{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table.
-This allows the numner of non-missing and the mean to be displayed on separate lines.
-The \texttt{"age"} variable's label is updated to \texttt{"Patient Age"}.
-Default summary statistics for both continuous and categorical variables are updated using the \texttt{statistic=} argument. 
-\pkg{gtsummary} uses \CRANpkg{glue} to help interpret string literals. Glue embeds R expressions (functions or objects) in curly braces which are then evaluated and inserted into the argument string. 
-ADD SOMETHING ABOUT GLUE SYNTAX!
-The \texttt{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \texttt{missing = "no"}.
+For continuous variables, tables display one row of statistics per variable by default.
+This can be customized, and in the example below, the age variable is cast to \code{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table. 
+This allows the number of non-missing observations and the mean to be displayed on separate lines.
 
+In the example below, the \code{"age"} variable's label is updated to \code{"Patient Age"}.
+Default summary statistics for both continuous and categorical variables are updated using the \code{statistic=} argument. 
+\pkg{gtsummary} uses \CRANpkg{glue} syntax construct the statistics displayed in the table.
+Function names appearing in curly brackets will be replaced by the evaluated value.
+The \code{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \code{missing = "no"}.
+
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{summary_plus.png}
+  \includegraphics[scale=0.35]{summary_plus.png}
   \centering
@@
 Throughout the \pkg{gtsummary} package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
-In the example above, the label for the age variable was updated using \texttt{label = age $\sim$ "Patient Age"}---equivalently, \texttt{label = list(age $\sim$ "Patient Age")}.
-To select groups of variables, utilize the select helpers from the tidyselect \citep{tidyselect} and \pkg{gtsummary}.
-The \texttt{all\_continuous()} selector is a convenient way to select all continuous variables.
-In the example above it could have been used to change the summary statistics for all continuous variables---\code{all\_continuous() ~ c("\{N\_nonmiss\}", "\{mean\} (\{sd\})")}. 
-Similarly, users may utilize \texttt{all\_categorical()} (from \pkg{gtsummary}), or any of the tidyselect helpers used throughout the tidyverse \citep{tidyverse} package, such as \texttt{starts\_with()}, \texttt{contains()}, etc.
+In the example above, the label for the age variable was updated using \code{label = age $\sim$ "Patient Age"}---equivalently, \code{label = list(age $\sim$ "Patient Age")}.
+To select groups of variables, utilize the select helpers from the \pkg{tidyselect} and \pkg{gtsummary}.
+The \code{all\_continuous()} selector is a convenient way to select all continuous variables.
+In the example above it could have been used to change the summary statistics for all continuous variables---\code{all\_continuous() $\sim$ c("\{N\_nonmiss\}", "\{mean\} (\{sd\})")}. 
+Similarly, users may utilize \code{all\_categorical()} (from \pkg{gtsummary}), or any of the \pkg{tidyselect} helpers used throughout the \pkg{tidyverse} package, such as \code{starts\_with()}, \code{contains()}, etc.
 
-In addition to basic summary statistics, the \pkg{gtsummary} package has several functions to add additional information or statistics to \texttt{tbl\_summary()} tables.
+In addition to summary statistics, the \pkg{gtsummary} package has several functions to add additional information or statistics to \code{tbl\_summary()} tables.
 
@@
 In the example below the number of non-missing observations is reported for each variable, as well as a p-value comparing the values between the treatment.
-Default statistical tests are chosen based on data type, but test type can be customized within the \texttt{add\_p()} function.
-P-value formatting can be adjusted using the \texttt{pvalue\_fun=} argument which accepts both a proper function, as well the formula shortcut notation used throughout tidyverse packages.
+Default statistical tests are chosen based on data type, and the statistical test performed can be customized in the \code{add\_p()} function.
+P-value formatting can be adjusted using the \code{pvalue\_fun=} argument which accepts both a proper function, as well the formula shortcut notation used throughout the \pkg{tidyverse} packages.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{summary_plus_plus.png}
+  \includegraphics[scale=0.35]{summary_plus_plus.png}
   \centering
@@
 
-\subsection{\texorpdfstring{\texttt{tbl\_svysummary()}}{tbl\_svysummary()}}
+\subsection{\texorpdfstring{\code{tbl\_svysummary()}}{tbl\_svysummary()}}
 
-The \texttt{tbl\_svysummary()} function is very similar to \texttt{tbl\_summary()} except a survey object is supplied rather than a data frame.
+The \code{tbl\_svysummary()} function is analagous to \code{tbl\_summary()} except an \CRANpkg{survey} object is supplied rather than a data frame.
+The summary statistics presented take into account the survey weights, as do any p-values presented.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{svysummary.png}
+  \includegraphics[scale=0.35]{svysummary.png}
   \centering
@@
 
-\subsection{\texorpdfstring{\texttt{tbl\_cross()}}{tbl\_cross()}}
+\subsection{\texorpdfstring{\code{tbl\_cross()}}{tbl\_cross()}}
 
-The \texttt{tbl\_cross()} function is a wrapper for \texttt{tbl\_summary()} that creates a simple, publication-ready cross tabulation table.
-The syntax is consistent with\texttt{tbl\_summary()} and \texttt{tbl\_svyummary()}
+The \code{tbl\_cross()} function is a wrapper for \code{tbl\_summary()} and creates a simple, publication-ready cross tabulation.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{cross.png}
+  \includegraphics[scale=0.35]{cross.png}
   \centering
@@
 
-\subsection{\texorpdfstring{\texttt{tbl\_survfit()}}{tbl\_survfit()}}
+\subsection{\texorpdfstring{\code{tbl\_survfit()}}{tbl\_survfit()}}
 
-The \texttt{tbl\_survfit()} function parses and tabulates \texttt{survival::survfit()} objects presenting survival percentile estimates and time point-specific survival probability estimates.
+The \code{tbl\_survfit()} function parses and tabulates \code{survival::survfit()} objects presenting survival percentile estimates and survival probabilities at specified times.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{survfit.png}
+  \includegraphics[scale=0.35]{survfit.png}
   \centering
@@
 
-The \pkg{gtsummary} package utilizes the gt package \citep{gt} to print the summary tables.
-The gt package exports approximately one hundred functions to customize and style tables.
-When you need to add additional details or styling not available within \pkg{gtsummary}, use the \texttt{as\_gt()} to convert the \pkg{gtsummary} object to gt and continue customization.
+The \pkg{gtsummary} package utilizes the \pkg{gt} package to print the summary tables.
+The \pkg{gt} package exports approximately one hundred functions to customize and style tables.
+When you need to add additional details or styling not available within \pkg{gtsummary}, use the \code{as\_gt()} to convert the \pkg{gtsummary} object to gt and continue customization.
 
-The example below includes customization using both gtsummary and gt functions.
-The gtsummary functions are utilized to bold p-values below a threshold, bold the variable labels,  update the column header, and add a spanning header
+The example below includes customization using both \pkg{gtsummary} and \pkg{gt} functions.
+The \pkg{gtsummary} functions are utilized to bold p-values below a threshold, bold the variable labels,  update the column header, and add a spanning header.
+Additional \pkg{gt} customization was utilized to add table captions and source notes.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{custom.png}
+  \includegraphics[scale=0.35]{custom.png}
   \centering
@@
 Regression modelling is one of the most common tools of medical researchers.
-The \pkg{gtsummary} has two functions to assist researchers and analysts prepare tabular summaries of regression models: \texttt{tbl\_regression()} and \texttt{tbl\_uvregression()}.
+The \pkg{gtsummary} has two functions to assist researchers and analysts prepare tabular summaries of regression models: \code{tbl\_regression()} and \code{tbl\_uvregression()}.
 
-\subsection{\texorpdfstring{\texttt{tbl\_regression()}}{tbl\_regression()}}
+\subsection{\texorpdfstring{\code{tbl\_regression()}}{tbl\_regression()}}
 
-The \texttt{tbl\_regression()} function takes a regression model object in R and returns a formatted table of regression model results. 
-Like \texttt{tbl\_summary()}, \texttt{tbl\_regression()} creates highly customizable analytic tables with sensible defaults.
+The \code{tbl\_regression()} function takes a regression model object in R and returns a formatted table of regression model results. 
+Like \code{tbl\_summary()}, \code{tbl\_regression()} creates highly customizable analytic tables with sensible defaults.
 Common regression models, such as logistic regression and Cox proportional hazards regression, are automatically identified and the tables headers are pre-filled with appropriate column headers (i.e.
@@
 
-In the example below, the logistic regression model is summarized with \texttt{tbl\_regression()}.
+In the example below, the logistic regression model is summarized with \code{tbl\_regression()}.
 Note that a reference row for Grade I has been added, and the variable labels have been carried through into the table.
-Using \texttt{exponentiate = TRUE}  weexponentiate the regression coefficients, yielding the odds ratios.
-The helper function \texttt{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
+Using \code{exponentiate = TRUE}  we exponentiate the regression coefficients, yielding the odds ratios.
+The helper function \code{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{regression.png}
+  \includegraphics[scale=0.35]{regression.png}
   \centering
@@
 
-The \texttt{tbl\_regression()} function leverages the huge effort behind the broom package's \texttt{tidy()} function \citep{broom} to perform the initial formatting of the regression object.
-Because \texttt{tbl\_regression()} utilizes \texttt{broom::tidy()}, there are many model types that are supported out of the box, such as \texttt{lm()}, \texttt{glm()}, \texttt{lme4::lmer()}, \texttt{lme4::glmer()}, \texttt{geepack::geeglm()}, \texttt{survival::coxph()}, \texttt{survival::survreg()}, \texttt{survival::clogit()}, \texttt{nnet::multinom()}, \texttt{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
-The \pkg{gtsummary} package also contains a handful of custom tidiers to report standardized coefficients and bootstrapped confidence intervals.
+The \code{tbl\_regression()} function leverages the huge effort behind the broom package's \code{tidy()} function \citep{broom} to perform the initial formatting of the regression object.
+Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
+The \pkg{gtsummary} package also contains built in tidiers to report standardized coefficients and bootstrapped confidence intervals, as well as the ability to provide your own custom tidiers.
 
-\subsection{\texorpdfstring{\texttt{tbl\_uvregression()}}{tbl\_uvregression()}}
+\subsection{\texorpdfstring{\code{tbl\_uvregression()}}{tbl\_uvregression()}}
 
-The \texttt{tbl\_uvregression()} function is a wrapper for \texttt{tbl\_regression()} that is useful when you need a series of univariate regression models.
-The user passes a data frame to \texttt{tbl\_uvregression()}, indicates what the outcome is, what regression model to run, and the function will return a formatted table of stacked univarate regression models.
+The \code{tbl\_uvregression()} function is a wrapper for \code{tbl\_regression()} that is useful when you need a series of univariate regression models.
+The user passes a data frame to \code{tbl\_uvregression()}, indicates what the outcome is, what regression model to run, and the function will return a formatted table of stacked univariate regression models.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{uvregression.png}
+  \includegraphics[scale=0.35]{uvregression.png}
   \centering
@@
 We often need to report the results from a table in the text of an R markdown report.
-The\texttt{inline\_text()}  function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
+The\code{inline\_text()}  function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
 
@@
 Typically, the odds ratio, confidence interval, and p-value would be hard-coded into a report, which can lead to reproducibility issues if the data is updated and the hard-coded statistics are not amended.
-A simple call to the \texttt{inline\_text()} function will dynamically add the model results to a an Rmarkdown report.
+A simple call to the \code{inline\_text()} function will dynamically add the model results to a an Rmarkdown report.
 
 \begin{quote}
-The odds ratio for age was \texttt{\textasciigrave{}r\ inline\_text(uvreg,\ variable\ =\ age)\textasciigrave{}}.
+The odds ratio for age was \code{\textasciigrave{}r\ inline\_text(uvreg,\ variable\ =\ age)\textasciigrave{}}.
 \end{quote}
@@
 
-The default pattern to display for a regression table is \texttt{"\{estimate\} (\{conf.level*100\}\% CI \{conf.low\}, \{conf.high\}; \{p.value\})"} (again using glue syntax), and can be modified with the \texttt{inline\_text(pattern=)} argument.
+The default pattern to display for a regression table is \code{"\{estimate\} (\{conf.level*100\}\% CI \{conf.low\}, \{conf.high\}; \{p.value\})"} (again using glue syntax), and can be modified with the \code{inline\_text(pattern=)} argument.
 
@@
 For example in cancer research, we often report on models about both a tumor's response to a treatment, and a treatment efficacy in reducing the risk of death.
-These two models are often reported side-by-side, and it's simple to construct this table using  the \texttt{tbl\_merge()} function.
-First, build a table for each regression model using \texttt{tbl\_regression()}, then merge the two tables with \texttt{tbl\_merge()}.
+These two models are often reported side-by-side, and it's simple to construct this table using  the \code{tbl\_merge()} function.
+First, build a table for each regression model using \code{tbl\_regression()}, then merge the two tables with \code{tbl\_merge()}.
 Any number of \pkg{gtsummary} tables can be merged with this function.
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{merge.png}
+  \includegraphics[scale=0.35]{merge.png}
   \centering
@@
 
-Similarly, any number of \pkg{gtsummary} tables may be stacked on top of one another using the \texttt{tbl\_stack()} function.
+Similarly, any number of \pkg{gtsummary} tables may be stacked on top of one another using the \code{tbl\_stack()} function.
 
@@
 The default styling (e.g.
-statistics displayed in \texttt{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
+statistics displayed in \code{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
 However, you'll likely submit to another journal, or your personal preferences differ from the defaults.
@@
 With themes, users can control default settings for existing functions (e.g.
-always present means instead of medians in \texttt{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
+always present means instead of medians in \code{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
 Several themes are available in the package that follow various journals' reporting guidelines, reduce cell padding, and recently language themes were introduced and the output of \pkg{gtsummary} tables can be translated into more than 10 languages.
 
-For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \texttt{"lb to ub"} instead of \texttt{"lb, ub"} (ampong other modifications).
+For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"} (among other modifications).
 
+\newpage
 \begin{example}
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.28]{nejm.png}
+  \includegraphics[scale=0.35]{nejm.png}
   \centering
@@
 Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
-The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \texttt{knitr::kable()}, \pkg{gtsummary}, \pkg{huxtable} and more.
-While \pkg{gt} is used as default for most outputs, you can easily convert to using print engine of choice using several conversion helper functions provided in the package (e.g.
-\texttt{as\_kable()}).
-The package is designed to interact with these print engines behind the scenes to reduce burden on users and you generally only need to be aware of them if you want to add advanced customizations (eg.
-example X with \texttt{tab\_header}).
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()}, \pkg{flextable}, \pkg{huxtable} and more.
+While \pkg{gt} is used as default for most outputs, you can easily convert to using your print engine of choice using several conversion helper functions provided in the package (e.g.
+\code{as\_kable()}).
+The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations (eg.
+example X with \code{tab\_header}).
+The default print engine can be systematically set using \pkg{gtsummary} themes.
 

### 2020-12-18 RJtemplate_gtsummary_submission.tex

@@
 
-The \code{tbl\_regression()} function leverages the huge effort behind the broom package's \code{tidy()} function \citep{broom} to perform the initial formatting of the regression object.
-Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the mice package[add ref]and more.
-The \pkg{gtsummary} package also contains built in tidiers to report standardized coefficients and bootstrapped confidence intervals, as well as the ability to provide your own custom tidiers.
+The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} package's \code{tidy()} function to perform the initial formatting of the regression object \citep{broom}.
+Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, and models built with the mice package[add ref]and more.
+The \pkg{gtsummary} package also contains built in tidying functions to report standardized coefficients and bootstrapped confidence intervals.
+We may pass a custom tidier as well.
+This is helpful when need to present non-standard modifications to your model results, such as reporting Wald confidence intervals, or presenting results with modified variance-covariance regression coefficients.
 
@@
 We often need to report the results from a table in the text of an R markdown report.
-The\code{inline\_text()}  function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
+The \code{inline\_text()}  function reports statistics from \pkg{gtsummary} tables inline in an R markdown document.
 
@@
 Typically, the odds ratio, confidence interval, and p-value would be hard-coded into a report, which can lead to reproducibility issues if the data is updated and the hard-coded statistics are not amended.
-A simple call to the \code{inline\_text()} function will dynamically add the model results to a an Rmarkdown report.
+A simple call to the \code{inline\_text()} function will dynamically add the model results to an RMarkdown report.
 
@@
 
-The \pkg{gtsummary} tables shown above are often ready for publication as they are, however it is common that higher quality tables can be constructed by merging or stacking simple tables.
-For example in cancer research, we often report on models about both a tumor's response to a treatment, and a treatment efficacy in reducing the risk of death.
-These two models are often reported side-by-side, and it's simple to construct this table using  the \code{tbl\_merge()} function.
+The \pkg{gtsummary} tables shown above are often ready for publication as they are; however, it is common that more complex tables need to be constructed, which can be achieved by merging or stacking \pkg{gtsummary} tables.
+For example in cancer research, we often report two model for predicting a tumor's response to treatment and risk of death.
+These two models are often reported side-by-side, and it's simple to construct using \code{tbl\_merge()}.
 First, build a table for each regression model using \code{tbl\_regression()}, then merge the two tables with \code{tbl\_merge()}.
@@
 
-Similarly, any number of \pkg{gtsummary} tables may be stacked on top of one another using the \code{tbl\_stack()} function.
+Similarly, any number of \pkg{gtsummary} tables may be stacked using the \code{tbl\_stack()} function.
 
@@
 
+We love themes.
 The default styling (e.g.
@@
 always present means instead of medians in \code{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
-Several themes are available in the package that follow various journals' reporting guidelines, reduce cell padding, and recently language themes were introduced and the output of \pkg{gtsummary} tables can be translated into more than 10 languages.
+Several themes are available to follow various journals' reporting guidelines, reduce cell padding and font size, and recently language themes were introduced translating \pkg{gtsummary} tables to more than 10 languages.
 

### 2020-12-18 RJtemplate_gtsummary_submission.tex
overleaf

@@
 % TO DO
-% - change all code refs from \code{} to \code{}
-% - update all references to packages to use \CRANpkg{} and \pkg{}
-% - scale the images appropriately (they are too small now)
+% - add/remove the page breaks as needed (when it's ready to be submitted)
 
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{summary_basic.png}
+  \includegraphics[scale=0.49]{summary_basic.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{summary_plus.png}
+  \includegraphics[scale=0.49]{summary_plus.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{summary_plus_plus.png}
+  \includegraphics[scale=0.49]{summary_plus_plus.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{svysummary.png}
+  \includegraphics[scale=0.49]{svysummary.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{cross.png}
+  \includegraphics[scale=0.49]{cross.png}
   \centering
@@
 
-% TODO: I do not understand why this isn't being placed below the code! https://www.overleaf.com/learn/latex/Positioning_of_Figures
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{survfit.png}
+  \includegraphics[scale=0.49]{survfit.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{custom.png}
+  \includegraphics[scale=0.49]{custom.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{regression.png}
+  \includegraphics[scale=0.49]{regression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{uvregression.png}
+  \includegraphics[scale=0.49]{uvregression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{merge.png}
+  \includegraphics[scale=0.49]{merge.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.35]{nejm.png}
+  \includegraphics[scale=0.49]{nejm.png}
   \centering
@@
 
-A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font. 
+A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font and reduce default cell padding. 
 Themes are an evolving feature and we welcome additions of new journals' reporting guidelines and other themes useful to users.
@@
 Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
-The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()}, \pkg{flextable}, \pkg{huxtable} and more.
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()}, \pkg{flextable}, \pkg{huxtable} and \pkg{kableExtra}.
 While \pkg{gt} is used as default for most outputs, you can easily convert to using your print engine of choice using several conversion helper functions provided in the package (e.g.
@@
 \section{Summary}
+
+The functions in the \pkg{gtsummary} package were designed to reduce the burden of reporting, and to work together to easily construct both simple and complex tables.
+It is our hope that the user-friendly syntax and publication-ready tables will aid analysts in preparing reproducible and high-quality findings.
 

### 2020-12-21 RJtemplate_gtsummary_submission.tex

@@
 
-Table summaries are a fundamental tool in an analyst's toolbox that help us understand and communicate patterns in our data,
+Table summaries are a fundamental tool in an analyst's toolbox that help us understand and communicate patterns in our data.
 The ability to easily create and export polished and reproducible tables is essential.
 The \CRANpkg{gtsummary} \citep{gtsummary} package provides an elegant and flexible framework to create publication-ready analytical and summary tables in R.
-This package works to gap the loop between a reproducible RMarkdown report and the final report.
+This package works to close the gap between a reproducible RMarkdown report and the final report.
 Specifically, \pkg{gtsummary} allows the us to fully customize and format summary tables with code, eliminating the need to modify any tables by hand after the table has been exported.
-Removing the need to modify tables after the table has been created, eliminates an error-prone step in our workflow and increases the reproducibility of our analyses and reports.
+Removing the need to modify tables after the table has been created eliminates an error-prone step in our workflow, and increases the reproducibility of our analyses and reports.
 
@@
 
-While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables,offers diverse features to customize content of tables to a high degree, and includes the ability to report statistics from the tables inline in an RMarkdown report.
+While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
+It also includes the ability to report statistics from the tables inline in an RMarkdown report.
 %For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
@@
 
-Along with descriptive summaries, \pkg{gtsummary} summarises statistical models, survey data, and build cross tabulations. 
+Along with descriptive summaries, \pkg{gtsummary} summarises statistical models, survey data, and builds cross tabulations. 
 After data are summarized in a table, \pkg{gtsummary} allows users to combine tables, either side-by-side (with \code{tbl\_merge()}) , or on top of each other (with \code{tbl\_stack()}). 
@@
 
-To showcase \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B as well as the outcomes of tumor response and death.
+To showcase \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B, as well as the outcomes of tumor response and death.
 Each variable in the data frame has been assigned an attribute label with the \pkg{labelled} package \citep{labelled}, e.g.
-\code{trial \%>\% set\_variable\_labels(age = "Age")}, that will be shown in the summary tables.
+\code{trial \%>\% set\_variable\_labels(age = "Age")} that will be shown in the summary tables.
 These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned the variable name would have been shown.
@@
 % code for basic tbl_summary()
-\newpage
+% \newpage
 \begin{example}
@@
 For continuous variables, tables display one row of statistics per variable by default.
-This can be customized, and in the example below, the age variable is cast to \code{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table. 
+This can be customized, and in the example below the age variable is cast to \code{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table. 
 This allows the number of non-missing observations and the mean to be displayed on separate lines.
@@
 Default summary statistics for both continuous and categorical variables are updated using the \code{statistic=} argument. 
-\pkg{gtsummary} uses \CRANpkg{glue} syntax construct the statistics displayed in the table.
+\pkg{gtsummary} uses \CRANpkg{glue} \citep{glue} syntax construct the statistics displayed in the table.
 Function names appearing in curly brackets will be replaced by the evaluated value.
-The \code{digits=} argument is used to increase the number of decimal places the statistics are rounded, and the missing row is omitted with \code{missing = "no"}.
+The \code{digits=} argument is used to increase the number of decimal places to which the statistics are rounded, and the missing row is omitted with \code{missing = "no"}.
 
-\newpage
+% \newpage
 \begin{example}
@@
 
-The \code{tbl\_svysummary()} function is analagous to \code{tbl\_summary()} except an \CRANpkg{survey} object is supplied rather than a data frame.
+The \code{tbl\_svysummary()} function is analagous to \code{tbl\_summary()} except an \CRANpkg{survey} \citep{survey} object is supplied rather than a data frame.
 The summary statistics presented take into account the survey weights, as do any p-values presented.
 
-\newpage
+% \newpage
 \begin{example}
@@
 
-\newpage
+% \newpage
 \begin{example}
@@
 These functions work with any table constructed with \pkg{gtsummary}.
-Most common uses are changing the column headers and footnotes.
+Most common uses are changing the column headers and footnotes, or modifying the look of tables through bolding and italicization.
 
@@
 The \pkg{gt} package exports approximately one hundred functions to customize and style tables.
-When you need to add additional details or styling not available within \pkg{gtsummary}, use the \code{as\_gt()} to convert the \pkg{gtsummary} object to gt and continue customization.
+When you need to add additional details or styling not available within \pkg{gtsummary}, use the \code{as\_gt()} function to convert the \pkg{gtsummary} object to gt and continue customization.
 
@@
 \end{example}
-\begin{figure}[h!]
+\begin{widefigure}[h!]
   \includegraphics[scale=0.49]{custom.png}
   \centering
-\end{figure}
+\end{widefigure}
 
@@
 
-Regression modelling is one of the most common tools of medical researchers.
-The \pkg{gtsummary} has two functions to assist researchers and analysts prepare tabular summaries of regression models: \code{tbl\_regression()} and \code{tbl\_uvregression()}.
+Regression modelling is one of the most common tools of medical research.
+The \pkg{gtsummary} package has two functions to help analysts prepare tabular summaries of regression models: \code{tbl\_regression()} and \code{tbl\_uvregression()}.
 
@@
 Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, and models built with the mice package[add ref]and more.
-The \pkg{gtsummary} package also contains built in tidying functions to report standardized coefficients and bootstrapped confidence intervals.
-We may pass a custom tidier as well.
-This is helpful when need to present non-standard modifications to your model results, such as reporting Wald confidence intervals, or presenting results with modified variance-covariance regression coefficients.
+The \pkg{gtsummary} package also contains built-in tidying functions to report standardized coefficients and bootstrapped confidence intervals.
+A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals, or results with modified variance-covariance regression coefficients.
 
@@
 
-\newpage
 \begin{example}
@@
 \end{example}
-
+\newpage
 \begin{figure}[h!]
@@
 
-The \pkg{gtsummary} tables shown above are often ready for publication as they are; however, it is common that more complex tables need to be constructed, which can be achieved by merging or stacking \pkg{gtsummary} tables.
+The \pkg{gtsummary} tables shown above are often ready for publication as they are; however, it is common that more complex tables need to be constructed.
+This can be achieved by merging or stacking \pkg{gtsummary} tables using the \code{tbl\_merge()} and \code{tbl\_stack()} functions.
 For example in cancer research, we often report two model for predicting a tumor's response to treatment and risk of death.
@@
 \end{example}
+\newpage
 
@@
 
-A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font and reduce default cell padding. 
-Themes are an evolving feature and we welcome additions of new journals' reporting guidelines and other themes useful to users.
+A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font and reduce the default cell padding. 
+Themes are an evolving feature and we welcome additions of new journals' reporting guidelines or other themes useful to users.
 A full glossary of customizable theme elements is available in the package's themes vignette (\url{http://www.danieldsjoberg.com/gtsummary/articles/themes.html}).
@@
 Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
-The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()}, \pkg{flextable}, \pkg{huxtable} and \pkg{kableExtra}.
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
 While \pkg{gt} is used as default for most outputs, you can easily convert to using your print engine of choice using several conversion helper functions provided in the package (e.g.
 \code{as\_kable()}).
-The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations (eg.
-example X with \code{tab\_header}).
+The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations.
 The default print engine can be systematically set using \pkg{gtsummary} themes.

### 2020-12-21 gtsummary_submission.tex
structure updates to folders

--- 2020-12-21 RJtemplate_gtsummary_submission.tex
@@
 Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
-The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{kable}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
 While \pkg{gt} is used as default for most outputs, you can easily convert to using your print engine of choice using several conversion helper functions provided in the package (e.g.

### 2020-12-21 tex_tables/tbl_trial_desc.tex
moved tables in to subfolder

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llll}
\caption{\label{tab:}Example data frame, \texttt{trial}}\\
\toprule
colname & label & class & values \\ 
\midrule
\texttt{trt} & Chemotherapy Treatment & character & \texttt{Drug A}, \texttt{Drug B} \\ 
\texttt{age} & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17} , ... \\ 
\texttt{marker} & Marker Level (ng/mL) & numeric & \texttt{0.003}, \texttt{0.005}, \texttt{0.013}, \texttt{0.015} , ... \\ 
\texttt{stage} & T Stage & factor & \texttt{T1}, \texttt{T2}, \texttt{T3}, \texttt{T4} \\ 
\texttt{grade} & Grade & factor & \texttt{I}, \texttt{II}, \texttt{III} \\ 
\texttt{response} & Tumor Response & integer & \texttt{0}, \texttt{1} \\ 
\texttt{death} & Patient Died & integer & \texttt{0}, \texttt{1} \\ 
\texttt{ttdeath} & Months to Death/Censor & numeric & \texttt{3.53}, \texttt{5.33}, \texttt{6.32}, \texttt{7.27} , ... \\ 
\bottomrule
\end{longtable}

### 2021-03-23 gtsummary_review_2021-03-23.Rmd
first pass at reviewer comments

---
title: "gtsummary Reviewer Responses"
author: "Dan Sjoberg, Karissa Whiting, Mike Curry"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: sandstone
    toc: yes
    toc_float: yes
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gtsummary)
library(gt)

```

Thank you for your reviews and for your consideration of our paper. We believe we have adequately addressed the reviewer's issues and have responded to the comments below.

# Reviewer 1

1. The authors didn’t provide any background review. Generating summary stats tables is such a common task that many people have tried to solve this problem before. I hope the authors could provide a background review of at least a few of the following packages

> We agree that a short section in the paper mentioning capabilities of other similar table making packages will help contextualize gtsummary's value. We have added this in the revised paper.

2. It seems that gtsummary has very limited support on generating LaTeX tables....It is okay to have limited support on some features but in terms of article writing, I would like to see the author mention it as a limitation in the discussion section.

3. The package is well written with solid documentation. I only have one question here. In `tbl_summary`, why let it have its own `by` argument instead of using `dplyr`’s `group_by`?

> One important difference between `dplyr`’s `group_by` and `tbl_summary`'s `by` argument is that you can group by many variables using `group_by`, but only one variable using `by`. Since `tbl_summary()` cannot support all the uses of `group_by`, we wanted to make a clear distinction that these have different functionalities.

# Reviewer 2

1. The article is very heavy in tables, which is understandable given the nature of the package. But maybe there is a way to rationalize? For instance, the `tbl_svysummary_1` table is essentially identical to the previous one. Do we need to see it in such a short exposition, or is it sufficient to discuss survey support?

> We have discovered many of gtsummary's most active users require support for survey data, so would like to keep this example in, unless the editor feels otherwise.

2. I’m not sure I understand the motivation for the table customization functions (e.g., bold_levels). Isn’t the whole point of building a package on top of gt that we can leverage its functions directly?

> While the package allows users to easily leverage gt customization, Knowledge of gt isn’t necessary to use the package. We’ve included some helper functions that make the most commonly needed customizations readily available to users who may not be familiar with {gt}, or may want to use any of the other print engines. Having the customization functions for these most common everyday customization tasks allows the users to make these customizations regardless of the print engine or output format they want to use, and customizations remain reproducible, even if they decide to switch print engine or output format.

> Most importantly, these functions helps the average users who may not be familiar with the internals of the package get. For example, to bold using {gt} functions, you have to have some knowledge of internal structure of gtsummary tables to select the correct columns and rows to customize.

```{r}
# include example?
```

3. Package Testing

> {gtsummary} currently has over XX tests. As noted, many of these tests check for specific warnings/messages, however many of these are looking for specific targeted warnings/messages that also test functionality, and there are still a considerable amount that check for expected returned values, etc.

> We are hesitant to add additional tests due to runtime/computational load, but will audit current tests and aim to change over some expect_error(), expect_warning() tests to more targeted tests for specific output values in the future.

4. tbl_summary: the continuous/categorical cutoff seems to be 10. can this be changed?

> Although you can't currently adjust the default cutoff, tidyselect helper functions and the `type`  argument make it very easy to specify what type you want your variables to be.

5. tbl_summary: Is there a way to set the number of digits for all variables at the same time? This can be onerous when there are many variables.

> tbl_summary() works with tidyselect functions, so you can specify all with the following:

```{r}
trial %>% 
  select(age, marker, ttdeath) %>%
  tbl_summary(digits = everything() ~ 2)

```

6. tbl_summary: Can we give a manual sort order?

> We added frequency sorting and alpha numeric for convenience, but we suggest to users that any variable or dataframe transformation should be done before passing tables to tbl_summary().

> For variable levels you can use `forcats::fct_relevel()`

```{r}

```

> For variable order you can use use `dplyr::select()`

```{r}

```

7. tbl_summary: Is it possible to use custom/arbitrary functions in the statistic argument, or are we limited to those explicitly defined by the package?

> TBD

8. tbl_regression: How do I manually reorder terms?

> TBD

### 2021-04-10 gtsummary_submission.tex
overleaf

--- 2020-12-21 RJtemplate_gtsummary_submission.tex
@@
 It also includes the ability to report statistics from the tables inline in an RMarkdown report.
+For example other summary statistics packages such as \CRANpkg{skimr}, \CRANpkg{summarytools}, and \CRANpkg{tableone} provide detailed descriptive statistics, but at least one major advantage of \CRANpkg{gtsummary} is the ability to also provide these same descriptive statistics that are publication ready in a scientific journal with little or no additional formatting.
+Additionally \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types so the user does not have to specify if a variable should be displayed with categorical or continuous summaries, unlike  \CRANpkg{tableone}.
+However, it is possible to override \CRANpkg{gtsummary} variable selection defaults if the users chooses too. 
 %For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
@@
 % adding table describing trial dataset
-\input{tbl_trial_desc.tex}
+\input{tex_tables/tbl_trial_desc.tex}
 
@@
 % code for tbl_summary() with arguments
-\input{tbl_tbl_summary_args.tex}
+\input{tex_tables/tbl_tbl_summary_args.tex}
 
@@
 % code for tbl_summary() with arguments
-\input{tbl_tbl_summary_family.tex}
+\input{tex_tables/tbl_tbl_summary_family.tex}
 
@@
 % adding table describing fns for customization
-\input{tbl_modify.tex}
+\input{tex_tables/tbl_modify.tex}
 

### 2021-04-10 reviewer_responses/gtsummary_review_2021-03-23.Rmd
updates

--- 2021-03-23 gtsummary_review_2021-03-23.Rmd
@@
 title: "gtsummary Reviewer Responses"
-author: "Dan Sjoberg, Karissa Whiting, Mike Curry"
+author: "Daniel Sjoberg, Karissa Whiting, Mike Curry"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 
-> We agree that a short section in the paper mentioning capabilities of other similar table making packages will help contextualize gtsummary's value. We have added this in the revised paper.
+> We agree that a section in the paper mentioning capabilities of other similar table making packages will help contextualize gtsummary's value. We have added this in the revised paper.
 
@@
 It seems that gtsummary has very limited support on generating LaTeX tables....It is okay to have limited support on some features but in terms of article writing, I would like to see the author mention it as a limitation in the discussion section.
+
+> The original submission did not make reference specically to the availability of LaTeX output.
+We've added text to the *Print Engine* section of the manuscript explicicty listing the output options, including LaTeX.
 
@@
 I’m not sure I understand the motivation for the table customization functions (e.g., bold_levels).
-Isn’t the whole point of building a package on top of gt that we can leverage its functions directly?
+Isn’t the whole point of building a package on top of {gt} that we can leverage its functions directly?
 
-> While the package allows users to easily leverage gt customization, Knowledge of gt isn’t necessary to use the package.
-We’ve included some helper functions that make the most commonly needed customizations readily available to users who may not be familiar with {gt}, or may want to use any of the other print engines.
-Having the customization functions for these most common everyday customization tasks allows the users to make these customizations regardless of the print engine or output format they want to use, and customizations remain reproducible, even if they decide to switch print engine or output format.
-
-> Most importantly, these functions helps the average users who may not be familiar with the internals of the package get.
-For example, to bold using {gt} functions, you have to have some knowledge of internal structure of gtsummary tables to select the correct columns and rows to customize.
-
-```{r}
-# include example?
-```
+> The reviewer raises a great point.
+One of the great features of gtsummary is teh accessabitily of every gt function to produce a highly customized table, when needed.
+We have included functions to assist in common formatting to make them more accessible to users.
+For example, to re-created the functionality of the `bold_labels()` function, the user would need to pass `as_gt() %>% gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_body(columns = gt::vars(label), rows = row_type == "label"))`.
+The helper function is much easier to read, and moreover, the gt version requires knowledge of hidden columns (e.g.
+`"row_type"`) to correctly select the label rows.
+Without these formatting helper functions, most passive users of the package would not have the knowledge to create the formatting on their own.
 
@@
 
-> {gtsummary} currently has over XX tests.
-As noted, many of these tests check for specific warnings/messages, however many of these are looking for specific targeted warnings/messages that also test functionality, and there are still a considerable amount that check for expected returned values, etc.
-
-> We are hesitant to add additional tests due to runtime/computational load, but will audit current tests and aim to change over some expect_error(), expect_warning() tests to more targeted tests for specific output values in the future.
+> As noted by the reviewer, many of the unit tests check for specific error/warnings/messages, which in some cases, may not be considered best practices.
+The {gtsummary} package, however, currently has 1100 unit tests with hundreds of tests that check for quite specific output.
+The sum total of the tests, in our opinion, provides great unit testing coverege testing a wide variety of options possible in the package.
 
@@
 
-> Although you can't currently adjust the default cutoff, tidyselect helper functions and the `type`  argument make it very easy to specify what type you want your variables to be.
+> Although you can't currently adjust the default cutoff, tidyselect helper functions and the `type=`  argument make it very easy to specify what type you want your variables to be.
 
@@
 
-> tbl_summary() works with tidyselect functions, so you can specify all with the following:
+> tbl_summary() works with all tidyselect functions (in addition to tidyselect-like functions exported by {gtsummary}), and you can specify all variables using the `everything()` selector.
+For example,
 
@@
   tbl_summary(digits = everything() ~ 2)
-
 ```
@@
 
-> We added frequency sorting and alpha numeric for convenience, but we suggest to users that any variable or dataframe transformation should be done before passing tables to tbl_summary().
-
-> For variable levels you can use `forcats::fct_relevel()`
-
-```{r}
-
-```
-
-> For variable order you can use use `dplyr::select()`
-
-```{r}
-
-```
+> The ordering of the variable in `tbl_summary()` aligns with the order the variables appear in the input data frame.
+To re-arrange the variables, use `select()` before passing the data frame to `tbl_summary()`, e.g.
+`trial %>% select(age, grade) %>% tbl_summary()`
 
@@
 
-> TBD
+> Yes!
+Any function in base R, another package, or user-defined can be used to summarize the continuos data in `tbl_summary()`.
+The only restriction is the first argument of the function must accept a vector, and return a scalar.
 
@@
 
-> TBD
+> The terms appear in the same order as regression function.
+The simplest way to reorder the terms to order them in, for example, the `lm()` call.
+If for some reason that is not possible, they can be re-ordered after the `tbl_regression()` table has been built using `modify_table_body()` with `dplyr::arrange()`.

### 2021-04-10 gtsummary_submission.R

@@ -257,15 +257,16 @@ tbl_merge_1 %>%
 
 # Themes -----------------------------------------------------------------------
 
-theme_gtsummary_journal("nejm")
+theme_gtsummary_journal("jama")
 
-tbl_nejm <-
+tbl_jama <-
   glm(response ~ age + grade, trial, family = binomial) %>%
   tbl_regression(exponentiate = TRUE)
 
-tbl_nejm %>%
+
+tbl_jama %>%
   as_gt() %>%
-  gtsave(file = "images/nejm.png")
+  gtsave(file = "images/jama.png")
 
 # Print Engines ----------------------------------------------------------------
 

### 2021-04-10 gtsummary_submission.tex
Updates from Overleaf

--- 2020-12-21 RJtemplate_gtsummary_submission.tex
@@
 
-While other R packages are available to clean and present summary statistics and regression results in statistical reports, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
-It also includes the ability to report statistics from the tables inline in an RMarkdown report.
-%For example with \CRANpkg{finalfit}, you have to manually specify a printing engine for the table to render.
-%Furthermore, \CRANpkg{tableone} reports on general descriptive statistics, but it is lacking in its ability to produce tables for regression modeling strategies in R.
+While other R packages are available to present data and regression model summary tables, such as such as \CRANpkg{skimr}, \CRANpkg{stargazer}, \CRANpkg{finalfit}, and \CRANpkg{tableone}, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
+The default \CRANpkg{gtsummary} table is suitable to be published in a scientific journal with little or no additional formatting.
+For example, \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types so the user does not have to specify if a variable should be displayed with categorical or continuous summaries, which yields the summary table with minimal code.
 
@@
 \begin{example}
-theme_gtsummary_journal("nejm")
+theme_gtsummary_journal("jama")
 
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{nejm.png}
+  \includegraphics[scale=0.49]{jama.png}
   \centering
@@
 \code{as\_kable()}).
+It's possible to get results in HTML, PDF (via \LaTeX), RTF, Microsoft Word, PowerPoint, Excel and others, utilizing the various print engines.
 The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations.
-The default print engine can be systematically set using \pkg{gtsummary} themes.
+A default print engine can be assigned using \pkg{gtsummary} themes.
 

### 2021-04-10 tbl_modify.tex
updates

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{ll}
\caption{\label{tab:} Functions to style and modify gtsummary tables}\\
\toprule
Function & Description \\ 
\midrule
\texttt{modify\_header()} & update column headers \\ 
\texttt{modify\_footnote()} & update column footnote \\ 
\texttt{modify\_spanning\_header()} & update spanning headers \\ 
\texttt{modify\_caption()} & update table caption/title \\ 
\texttt{bold\_labels()} & bold variable labels \\ 
\texttt{bold\_levels()} & bold variable levels \\ 
\texttt{italicize\_labels()} & italicize variable labels \\ 
\texttt{italicize\_levels()} & italicize variable levels \\ 
\texttt{bold\_p()} & bold significant p-values \\ 
\bottomrule
\end{longtable}

### 2021-04-10 gtsummary_submission.R

@@ -2,6 +2,7 @@ library(gtsummary)
 library(tidyverse)
 library(gt)
 
+reset_gtsummary_theme()
 # theme_gtsummary_compact()
 # set_gtsummary_theme(list(
 #   "as_gt-lst:addl_cmds" = list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino")))
@@ -165,6 +166,7 @@ tibble::tribble(
   "`modify_header()`",           "update column headers",
   "`modify_footnote()`",         "update column footnote",
   "`modify_spanning_header()`",  "update spanning headers",
+  "`modify_caption()`",          "update table caption/title",
   "`bold_labels()`",             "bold variable labels",
   "`bold_levels()`",             "bold variable levels",
   "`italicize_labels()`",        "italicize variable labels",

### 2021-04-10 reviewer_responses/gtsummary_review_2021-03-23.Rmd
updates

@@
     toc_float: yes
-    number_sections: true
 editor_options: 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
 
@@
 
-3. Package Testing
+3.
+The gtsummary developers use testthat as their testing suite, and codecov to measure and report their test coverage.
+On Github, they report a 93% test coverage which, at first glance, is very impressive.
+Unfortunately, I do not think that those tests follow best practices, and I am not convinced that the 93% tells us much.
+After looking at a few tests I did a quick grep and found that about 75% of expectations in the test folder were either expect_message, expect_error, or expect_warning (please correct me if this is wrong!).
+Most of the time, the tests only seem to make sure that no error is produced.
 

### 2021-04-10 gtsummary_submission.tex

@@
 In the example below, the logistic regression model is summarized with \code{tbl\_regression()}.
-Note that a reference row for Grade I has been added, and the variable labels have been carried through into the table.
+Note that a reference row for Grade has been added, and the variable labels have been carried through into the table.
 Using \code{exponentiate = TRUE}  we exponentiate the regression coefficients, yielding the odds ratios.
@@
 The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} package's \code{tidy()} function to perform the initial formatting of the regression object \citep{broom}.
-Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, and models built with the mice package[add ref]and more.
+Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and more.
 The \pkg{gtsummary} package also contains built-in tidying functions to report standardized coefficients and bootstrapped confidence intervals.
@@
 
-Reproducible reports are an important part of good analysis practices.
+Reproducible reports are an important part of good analytic practices.
 We often need to report the results from a table in the text of an R markdown report.
@@
 always present means instead of medians in \code{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
-Several themes are available to follow various journals' reporting guidelines, reduce cell padding and font size, and recently language themes were introduced translating \pkg{gtsummary} tables to more than 10 languages.
+Several themes are available to follow various journals' reporting guidelines, reduce cell padding and font size, and language themes to translate \pkg{gtsummary} tables to more than 14 languages.
 
-For example, using the theme for New England Journal of Medicine, large p-values are rounded to two decimal places and confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"} (among other modifications).
+For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence internal is displayed in the same column as the model coefficient.
 
@@
 The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
-While \pkg{gt} is used as default for most outputs, you can easily convert to using your print engine of choice using several conversion helper functions provided in the package (e.g.
-\code{as\_kable()}).
+While \pkg{gt} is used as default for most outputs, you can easily use your print engine of choice with the conversion helper functions provided in the package (e.g. \code{as\_flex\_table()}).
 It's possible to get results in HTML, PDF (via \LaTeX), RTF, Microsoft Word, PowerPoint, Excel and others, utilizing the various print engines.

### 2021-04-10 gtsummary_submission.R
tex table update

@@ -39,7 +39,7 @@ trial %>%
   as_latex() %>%
   as.character() %>%
   str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}Example data frame, \\texttt{trial}}\\\\\n\\toprule")) %>%
-  readr::write_lines("tbl_trial_desc.tex")
+  readr::write_lines("tex_tables/tbl_trial_desc.tex")
 
 ## `tbl_summary()` -------------------------------------------------------------
 
@@ -69,7 +69,7 @@ tibble::tribble(
   as_latex() %>%
   as.character() %>%
   str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl\\_summary()} function arguments}\\\\\n\\toprule")) %>%
-  readr::write_lines("tbl_tbl_summary_args.tex")
+  readr::write_lines("tex_tables/tbl_tbl_summary_args.tex")
 
 tbl_summary_2 <-
   trial %>%
@@ -104,7 +104,7 @@ tibble::tribble(
   as_latex() %>%
   as.character() %>%
   str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl\\_summary()} functions to add information}\\\\\n\\toprule")) %>%
-  readr::write_lines("tbl_tbl_summary_family.tex")
+  readr::write_lines("tex_tables/tbl_tbl_summary_family.tex")
 
 tbl_summary_3 <-
   trial %>%
@@ -179,7 +179,7 @@ tibble::tribble(
   as_latex() %>%
   as.character() %>%
   str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:} Functions to style and modify gtsummary tables}\\\\\n\\toprule")) %>%
-  readr::write_lines("tbl_modify.tex")
+  readr::write_lines("tex_tables/tbl_modify.tex")
 
 tbl_custom <-
   trial %>%

### 2021-04-13 gtsummary_submission.R
updates!

@@ -183,19 +183,18 @@ tibble::tribble(
 
 tbl_custom <-
   trial %>%
-  select(marker, grade, trt) %>%
-  tbl_summary(by = trt, missing = "no") %>%
+  select(marker, response, trt) %>%
+  tbl_summary(by = trt, missing = "no",
+              statistic = list(marker ~ "{mean} ({sd})")) %>%
+  add_difference() %>%
   add_overall() %>%
-  add_p() %>%
-  add_q() %>%
   add_stat_label() %>%
-  bold_p(t = 0.10) %>%
   bold_labels() %>%
   modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Received**") %>%
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Randomization Assignment**") %>%
   as_gt() %>%
   gt::tab_header(
-    title = gt::md("**Table 1. Patient Characteristics**"),
+    title = gt::md("**Table 1. Treatment Differences**"),
     subtitle = gt::md("_Highly Confidential_")
   ) %>%
   gt::tab_source_note("Data updated June 26, 2015")

### 2021-04-13 gtsummary_submission.tex
Updates from Overleaf

@@
 
-While other R packages are available to present data and regression model summary tables, such as such as \CRANpkg{skimr}, \CRANpkg{stargazer}, \CRANpkg{finalfit}, and \CRANpkg{tableone}, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
+While other R packages are available to present data and regression model summary tables, such as \CRANpkg{skimr}, \CRANpkg{stargazer}, \CRANpkg{finalfit}, and \CRANpkg{tableone}, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
 The default \CRANpkg{gtsummary} table is suitable to be published in a scientific journal with little or no additional formatting.
-For example, \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types so the user does not have to specify if a variable should be displayed with categorical or continuous summaries, which yields the summary table with minimal code.
+For example, \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types so the user does not have to specify if a variable should be displayed with categorical or continuous summaries, which yields summary tables with minimal code.
 
@@
 \code{trial \%>\% set\_variable\_labels(age = "Age")} that will be shown in the summary tables.
-These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned the variable name would have been shown.
+These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned, the variable name would have been shown.
 
@@
 trial %>%
-  select(age, grade, trt) %>%
-  tbl_summary(by = trt, missing = "no") %>%
+  select(marker, response, trt) %>%
+  tbl_summary(by = trt, missing = "no", 
+              statistic = list(marker ~ "{mean} ({sd})")) %>%
   add_overall() %>%
-  add_p() %>%
-  add_q() %>%
+  add_difference() %>%
   add_stat_label() %>%
-  bold_p(t = 0.10) %>%
   bold_labels() %>%
   modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c(stat_1, stat_2) ~ "**Treatment Received**") %>%
+  modify_spanning_header(c(stat_1, stat_2) ~ "**Randomization Assignment**") %>%
   as_gt() %>%
   gt::tab_header(
-    title = gt::md("**Table 1. Patient Characteristics**"),
+    title = gt::md("**Table 1. Treatment Differences**"),
     subtitle = gt::md("_Highly Confidential_")

### 2021-04-13 gtsummary_submission.R
updates

@@ -96,7 +96,7 @@ tibble::tribble(
   "`add_difference()`",  "add column for difference between two group, confidence interval, and p-value",
   "`add_stat_label()`",  "add label for the summary statistics shown in each row",
   "`add_stat()`",        "generic function to add a column with user-defined values",
-  "`add_q()`",           "add a column of q values to control for multiple comparisons"
+  "`add_q()`",           "add a column of q-values to control for multiple comparisons"
 ) %>%
   gt(caption = "`tbl_summary()` functions to add information") %>%
   fmt_markdown(everything()) %>%
@@ -112,7 +112,7 @@ tbl_summary_3 <-
   tbl_summary(by = trt, missing = "no") %>%
   add_p(test = all_continuous() ~ "t.test",
         pvalue_fun = ~style_pvalue(., digits = 2)) %>%
-  add_n()
+  add_overall()
 
 tbl_summary_3 %>%
   as_gt() %>%
@@ -187,10 +187,12 @@ tbl_custom <-
   tbl_summary(by = trt, missing = "no",
               statistic = list(marker ~ "{mean} ({sd})")) %>%
   add_difference() %>%
-  add_overall() %>%
+  add_n() %>%
   add_stat_label() %>%
   bold_labels() %>%
-  modify_header(label ~ "**Variable**") %>%
+  modify_header(
+    list(label ~ "**Variable**",
+         all_stat_cols() ~ "**{level}**")) %>%
   modify_spanning_header(c(stat_1, stat_2) ~ "**Randomization Assignment**") %>%
   as_gt() %>%
   gt::tab_header(

### 2021-04-13 gtsummary_submission.tex

@@
 
+\newpage
 % adding table describing trial dataset
@@
 
+\newpage
 % adding table describing fns for customization
@@
 The example below includes customization using both \pkg{gtsummary} and \pkg{gt} functions.
-The \pkg{gtsummary} functions are utilized to bold p-values below a threshold, bold the variable labels,  update the column header, and add a spanning header.
+The \pkg{gtsummary} functions are utilized to bold the variable labels,  update the column headers, and add a spanning header.
 Additional \pkg{gt} customization was utilized to add table captions and source notes.
@@
   select(marker, response, trt) %>%
-  tbl_summary(by = trt, missing = "no", 
+  tbl_summary(by = trt, 
+              missing = "no", 
               statistic = list(marker ~ "{mean} ({sd})")) %>%
@@
 
-For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence internal is displayed in the same column as the model coefficient.
+For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence internal is displayed in the same column as the model coefficients.
 

### 2021-04-13 gtsummary_submission.R
updates!

@@ -109,7 +109,7 @@ tibble::tribble(
 tbl_summary_3 <-
   trial %>%
   select(age, grade, response, trt) %>%
-  tbl_summary(by = trt, missing = "no") %>%
+  tbl_summary(by = trt) %>%
   add_p(test = all_continuous() ~ "t.test",
         pvalue_fun = ~style_pvalue(., digits = 2)) %>%
   add_overall()
@@ -193,7 +193,7 @@ tbl_custom <-
   modify_header(
     list(label ~ "**Variable**",
          all_stat_cols() ~ "**{level}**")) %>%
-  modify_spanning_header(c(stat_1, stat_2) ~ "**Randomization Assignment**") %>%
+  modify_spanning_header(all_stat_cols() ~ "**Randomization Assignment**") %>%
   as_gt() %>%
   gt::tab_header(
     title = gt::md("**Table 1. Treatment Differences**"),

### 2021-04-13 gtsummary_submission.R
Update gtsummary_submission.R

@@ -110,9 +110,9 @@ tbl_summary_3 <-
   trial %>%
   select(age, grade, response, trt) %>%
   tbl_summary(by = trt) %>%
+  add_overall() %>%
   add_p(test = all_continuous() ~ "t.test",
-        pvalue_fun = ~style_pvalue(., digits = 2)) %>%
-  add_overall()
+        pvalue_fun = ~style_pvalue(., digits = 2))
 
 tbl_summary_3 %>%
   as_gt() %>%

### 2021-04-13 gtsummary_submission.tex
Updates from Overleaf

@@
 
-In the example below the number of non-missing observations is reported for each variable, as well as a p-value comparing the values between the treatment.
+In the example below, descriptive statistics are shown by the treatment received and overall, as well as a p-value comparing the values between the treatments.
 Default statistical tests are chosen based on data type, and the statistical test performed can be customized in the \code{add\_p()} function.
@@
   select(age, grade, response, trt) %>%
-  tbl_summary(by = trt, missing = "no") %>%
-  add_n() %>%
+  tbl_summary(by = trt) %>%
+  add_overall() %>%
   add_p(test = all_continuous() ~ "t.test",
@@
 
-The example below includes customization using both \pkg{gtsummary} and \pkg{gt} functions.
+The example below is a common table reported in clinical trials and observational research where two treatments are compared.
+The treatment differences were added with the \code{add\_difference()} function.
+The table includes customization using both \pkg{gtsummary} and \pkg{gt} functions.
 The \pkg{gtsummary} functions are utilized to bold the variable labels,  update the column headers, and add a spanning header.
@@
   select(marker, response, trt) %>%
-  tbl_summary(by = trt, 
-              missing = "no", 
+  tbl_summary(by = trt, missing = "no",
               statistic = list(marker ~ "{mean} ({sd})")) %>%
-  add_overall() %>%
   add_difference() %>%
+  add_n() %>%
   add_stat_label() %>%
   bold_labels() %>%
-  modify_header(label ~ "**Variable**") %>%
-  modify_spanning_header(c(stat_1, stat_2) ~ "**Randomization Assignment**") %>%
+  modify_header(list(label ~ "**Variable**", all_stat_cols() ~ "**{level}**")) %>%
+  modify_spanning_header(all_stat_cols() ~ "**Randomization Assignment**") %>%
   as_gt() %>%
@@
 The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} package's \code{tidy()} function to perform the initial formatting of the regression object \citep{broom}.
-Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and more.
-The \pkg{gtsummary} package also contains built-in tidying functions to report standardized coefficients and bootstrapped confidence intervals.
+Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice} and many more.
 A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals, or results with modified variance-covariance regression coefficients.
@@
 This can be achieved by merging or stacking \pkg{gtsummary} tables using the \code{tbl\_merge()} and \code{tbl\_stack()} functions.
-For example in cancer research, we often report two model for predicting a tumor's response to treatment and risk of death.
+For example in cancer research, we often report models predicting a tumor's response to treatment and risk of death.
 These two models are often reported side-by-side, and it's simple to construct using \code{tbl\_merge()}.
@@
 The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations.
-A default print engine can be assigned using \pkg{gtsummary} themes.
 

### 2021-04-13 gtsummary_submission.tex
overleaf

@@
   select(marker, response, trt) %>%
-  tbl_summary(by = trt, missing = "no",
-              statistic = list(marker ~ "{mean} ({sd})")) %>%
+  tbl_summary(by = trt, 
+              missing = "no",
+              statistic = marker ~ "{mean} ({sd})") %>%
   add_difference() %>%

### 2021-04-13 reviewer_responses/gtsummary_review_2021-03-23.Rmd
small updates

@@
 > The reviewer raises a great point.
-One of the great features of gtsummary is teh accessabitily of every gt function to produce a highly customized table, when needed.
+One of the great features of gtsummary is the accessibility of every gt function to produce a highly customized table, when needed.
 We have included functions to assist in common formatting to make them more accessible to users.
@@
 To re-arrange the variables, use `select()` before passing the data frame to `tbl_summary()`, e.g.
-`trial %>% select(age, grade) %>% tbl_summary()`
+`trial %>% select(age, grade) %>% tbl_summary()`.
+Custom ordering of categorical variable levels can be done manually, or using forcats convenience functions (e.g.
+`forcats:fct_relevel()`) before passing data.
 
@@
 > Yes!
-Any function in base R, another package, or user-defined can be used to summarize the continuos data in `tbl_summary()`.
-The only restriction is the first argument of the function must accept a vector, and return a scalar.
+Any function in base R, another package, or user-defined can be used to summarize the continuous data in `tbl_summary()`.
+The only restriction is the first argument of the function must accept a vector and return a scalar.
 
@@
 
-> The terms appear in the same order as regression function.
-The simplest way to reorder the terms to order them in, for example, the `lm()` call.
+> The terms appear in the same order as in regression function.
+The simplest way to reorder the terms is to order them directly in call, for example, in the `lm()` call.
 If for some reason that is not possible, they can be re-ordered after the `tbl_regression()` table has been built using `modify_table_body()` with `dplyr::arrange()`.

### 2021-04-14 gtsummary_submission.tex
Updates from Overleaf
overleaf

@@
 
-Along with descriptive summaries, \pkg{gtsummary} summarises statistical models, survey data, and builds cross tabulations. 
+Along with descriptive summaries, \pkg{gtsummary} summarizes statistical models, survey data, and builds cross tabulations. 
 After data are summarized in a table, \pkg{gtsummary} allows users to combine tables, either side-by-side (with \code{tbl\_merge()}) , or on top of each other (with \code{tbl\_stack()}). 
-This table merging and stacking allows us to easily synthesize and compare output from several tables and share information in a compact format.
+The table merging and stacking abilities allows analysts to easily synthesize and compare output from several tables and share information in a compact format.
+All tables in this manuscript were created using \pkg{gtsummary} v1.4.0.
 
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{summary_basic.png}
+  \includegraphics[scale=0.73]{summary_basic.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{summary_plus.png}
+  \includegraphics[scale=0.73]{summary_plus.png}
   \centering
@@
 In the example above, the label for the age variable was updated using \code{label = age $\sim$ "Patient Age"}---equivalently, \code{label = list(age $\sim$ "Patient Age")}.
-To select groups of variables, utilize the select helpers from the \pkg{tidyselect} and \pkg{gtsummary}.
+To select groups of variables, utilize the select helpers from the \pkg{tidyselect} and \pkg{gtsummary} packages.
 The \code{all\_continuous()} selector is a convenient way to select all continuous variables.
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{summary_plus_plus.png}
+  \includegraphics[scale=0.73]{summary_plus_plus.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{svysummary.png}
+  \includegraphics[scale=0.73]{svysummary.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{cross.png}
+  \includegraphics[scale=0.73]{cross.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{survfit.png}
+  \includegraphics[scale=0.73]{survfit.png}
   \centering
@@
 \begin{widefigure}[h!]
-  \includegraphics[scale=0.49]{custom.png}
+  \includegraphics[scale=0.73]{custom.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{regression.png}
+  \includegraphics[scale=0.73]{regression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{uvregression.png}
+  \includegraphics[scale=0.73]{uvregression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{merge.png}
+  \includegraphics[scale=0.73]{merge.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.49]{jama.png}
+  \includegraphics[scale=0.73]{jama.png}
   \centering

### 2021-04-14 response_to_reviewers/gtsummary_review_responses_2021-03-23.Rmd
edits for resubmission

---
title: "gtsummary Reviewer Responses"
author: "Daniel Sjoberg, Karissa Whiting, Mike Curry"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: sandstone
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(gtsummary)
library(gt)
```

Thank you for your reviews and for your consideration of our paper. We believe we have adequately addressed the comments and have responded below.

# Reviewer 1

1. The authors didn’t provide any background review. Generating summary stats tables is such a common task that many people have tried to solve this problem before. I hope the authors could provide a background review of at least a few of the following packages

> We agree that a section in the paper mentioning capabilities of other similar table making packages will help contextualize gtsummary's value. We have added this in the revised paper.

2. It seems that gtsummary has very limited support on generating LaTeX tables....It is okay to have limited support on some features but in terms of article writing, I would like to see the author mention it as a limitation in the discussion section.

> The original submission did not make reference specically to the availability of LaTeX output. We've added text to the *Print Engine* section of the manuscript explicicty listing the output options, including LaTeX.

3. The package is well written with solid documentation. I only have one question here. In `tbl_summary`, why let it have its own `by` argument instead of using `dplyr`’s `group_by`?

> One important difference between `dplyr`’s `group_by` and `tbl_summary`'s `by` argument is that you can group by many variables using `group_by`, but only one variable using `by`. Since `tbl_summary()` cannot support all the uses of `group_by`, we wanted to make a clear distinction that these have different functionalities.

# Reviewer 2

1. The article is very heavy in tables, which is understandable given the nature of the package. But maybe there is a way to rationalize? For instance, the `tbl_svysummary_1` table is essentially identical to the previous one. Do we need to see it in such a short exposition, or is it sufficient to discuss survey support?

> We have discovered many of gtsummary's most active users require support for survey data, so would like to keep this example in, unless the editor feels otherwise.

2. I’m not sure I understand the motivation for the table customization functions (e.g., bold_levels). Isn’t the whole point of building a package on top of {gt} that we can leverage its functions directly?

> The reviewer raises a great point. One of the great features of gtsummary is the accessibility of every gt function to produce a highly customized table, when needed. We have included functions to assist in common formatting to make them more accessible to users. For example, to re-created the functionality of the `bold_labels()` function, the user would need to pass `as_gt() %>% gt::tab_style(style = gt::cell_text(weight = "bold"), locations = gt::cells_body(columns = gt::vars(label), rows = row_type == "label"))`. The helper function is much easier to read, and moreover, the gt version requires knowledge of hidden columns (e.g. `"row_type"`) to correctly select the label rows. Without these formatting helper functions, most passive users of the package would not have the knowledge to create the formatting on their own.

3. The gtsummary developers use testthat as their testing suite, and codecov to measure and report their test coverage. On Github, they report a 93% test coverage which, at first glance, is very impressive. Unfortunately, I do not think that those tests follow best practices, and I am not convinced that the 93% tells us much. After looking at a few tests I did a quick grep and found that about 75% of expectations in the test folder were either expect_message, expect_error, or expect_warning (please correct me if this is wrong!). Most of the time, the tests only seem to make sure that no error is produced.

> As noted by the reviewer, many of the unit tests check for specific error/warnings/messages, which in some cases, may not be considered best practices. The {gtsummary} package, however, currently has 1100 unit tests with hundreds of tests that check for quite specific output. The sum total of the tests, in our opinion, provides great unit testing coverege testing a wide variety of options possible in the package.

4. tbl_summary: the continuous/categorical cutoff seems to be 10. can this be changed?

> Although you can't currently adjust the default cutoff, tidyselect helper functions and the `type=`  argument make it very easy to specify what type you want your variables to be.

5. tbl_summary: Is there a way to set the number of digits for all variables at the same time? This can be onerous when there are many variables.

> tbl_summary() works with all tidyselect functions (in addition to tidyselect-like functions exported by {gtsummary}), and you can specify all variables using the `everything()` selector. For example,

```{r}
trial %>% 
  select(age, marker, ttdeath) %>%
  tbl_summary(digits = everything() ~ 2)
```

6. tbl_summary: Can we give a manual sort order?

> The ordering of the variable in `tbl_summary()` aligns with the order the variables appear in the input data frame. To re-arrange the variables, use `select()` before passing the data frame to `tbl_summary()`, e.g. `trial %>% select(age, grade) %>% tbl_summary()`. Custom ordering of categorical variable levels can be done manually, or using forcats convenience functions (e.g. `forcats:fct_relevel()`) before passing data.

7. tbl_summary: Is it possible to use custom/arbitrary functions in the statistic argument, or are we limited to those explicitly defined by the package?

> Yes! Any function in base R, another package, or user-defined can be used to summarize the continuous data in `tbl_summary()`. The only restriction is the first argument of the function must accept a vector and return a scalar.

8. tbl_regression: How do I manually reorder terms?

> The terms appear in the same order as in regression function. The simplest way to reorder the terms is to order them directly in call, for example, in the `lm()` call. If for some reason that is not possible, they can be re-ordered after the `tbl_regression()` table has been built using `modify_table_body()` with `dplyr::arrange()`.

### 2021-05-18 gtsummary_submission.tex

@@
 Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize regression models, and report statistics inline in RMarkdown reports.
-After identifying these basic structures of most tables presented in medical literature (and other fields), we created \pkg{gtsummary} to ease creation of fully-formatted, ready to publish tables.
+After identifying these basic structures of most tables presented in medical literature (and other fields), we wrote \pkg{gtsummary} to ease creation of fully-formatted, ready to publish tables.
 
@@
 HTML, PDF, Word, or RTF).
-Furthermore, \pkg{gtsummary} functions are designed to work within a "tidy" framework, utilizing the \CRANpkg{magrittr} pipe operator and \CRANpkg{tidyselect} \citep{tidyselect} functions used throughout the \CRANpkg{tidyverse} \citep{tidyverse}.
+Furthermore, \pkg{gtsummary} functions are designed to work within a "tidy" framework, utilizing the \CRANpkg{magrittr} \citep{magrittr} pipe operator and \CRANpkg{tidyselect} \citep{tidyselect} functions used throughout the \CRANpkg{tidyverse} \citep{tidyverse}.
 

### 2021-05-18 gtsummary_submission.tex
overleaf

@@
 Each variable in the data frame has been assigned an attribute label with the \pkg{labelled} package \citep{labelled}, e.g.
-\code{trial \%>\% set\_variable\_labels(age = "Age")} that will be shown in the summary tables.
+\code{trial \%>\% set\_variable\_labels(age = "Age")}, that will be shown in the summary tables.
 These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned, the variable name would have been shown.
@@
 
-The default output from \code{tbl\_summary()} is meant to be publication ready.
+The default output from \code{tbl\_summary()} is meant to be publication-ready.
 The \code{tbl\_summary()} function can take, at minimum, a data frame as the only input, and returns descriptive statistics for each column in the data frame.
@@
 For example, variables coded as \code{0/1}, \code{TRUE/FALSE}, and \code{Yes/No} are presented dichotomously.
-Additionally, \code{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are automatically utilized.
+Additionally, \code{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are utilized.
 
@@
 Default summary statistics for both continuous and categorical variables are updated using the \code{statistic=} argument. 
-\pkg{gtsummary} uses \CRANpkg{glue} \citep{glue} syntax construct the statistics displayed in the table.
+\pkg{gtsummary} uses \CRANpkg{glue} \citep{glue} syntax to construct the statistics displayed in the table.
 Function names appearing in curly brackets will be replaced by the evaluated value.
@@
 
-The \code{tbl\_svysummary()} function is analagous to \code{tbl\_summary()} except an \CRANpkg{survey} \citep{survey} object is supplied rather than a data frame.
+The \code{tbl\_svysummary()} function is analogous to \code{tbl\_summary()} except a \CRANpkg{survey} \citep{survey} object is supplied rather than a data frame.
 The summary statistics presented take into account the survey weights, as do any p-values presented.
@@
 The \pkg{gt} package exports approximately one hundred functions to customize and style tables.
-When you need to add additional details or styling not available within \pkg{gtsummary}, use the \code{as\_gt()} function to convert the \pkg{gtsummary} object to gt and continue customization.
+When you need to add additional details or styling not available within \pkg{gtsummary}, use the \code{as\_gt()} function to convert the \pkg{gtsummary} object to \pkg{gt} and continue customization.
 
@@
 Note that a reference row for Grade has been added, and the variable labels have been carried through into the table.
-Using \code{exponentiate = TRUE}  we exponentiate the regression coefficients, yielding the odds ratios.
+Using \code{exponentiate = TRUE}, we exponentiate the regression coefficients, yielding the odds ratios.
 The helper function \code{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
@@
 Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice} and many more.
-A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals, or results with modified variance-covariance regression coefficients.
+A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals, or results with modified variance-covariance standard errors.
 
@@
 This can be achieved by merging or stacking \pkg{gtsummary} tables using the \code{tbl\_merge()} and \code{tbl\_stack()} functions.
-For example in cancer research, we often report models predicting a tumor's response to treatment and risk of death.
+For example, in cancer research we often report models predicting a tumor's response to treatment and risk of death.
 These two models are often reported side-by-side, and it's simple to construct using \code{tbl\_merge()}.
@@
 
-Tables printed with \pkg{gtsummary} can be seamlessly integrated into R markdown documents and knitted into various output types using a number of print engines.
+Tables printed with \pkg{gtsummary} can be seamlessly integrated into RMarkdown documents and knitted into various output types using a number of print engines.
 The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.

### 2021-06-21 gtsummary_submission.R
image related updates

@@ -2,22 +2,22 @@ library(gtsummary)
 library(tidyverse)
 library(gt)
 
+# set gtsummary theme for the R journal ----------------------------------------
 reset_gtsummary_theme()
-# theme_gtsummary_compact()
-# set_gtsummary_theme(list(
-#   "as_gt-lst:addl_cmds" = list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino")))
-# ))
-
 my_r_journal_theme <- theme_gtsummary_compact(set_theme = FALSE)
 my_r_journal_theme$`as_gt-lst:addl_cmds` <-
   my_r_journal_theme$`as_gt-lst:addl_cmds` %>%
   c(list("cols_hide" = rlang::expr(gt::tab_options(table.font.names = "palatino"))))
-
 set_gtsummary_theme(my_r_journal_theme)
+
+# set to TRUE to re-render all tables (gt images and latex tables)
+# if you're at the R Journal, please leave this as FALSE
+save_tables <- FALSE
 
 # Data Summaries ---------------------------------------------------------------
 
-trial %>%
+if (save_tables)
+  trial %>%
   imap_dfr(
     ~tibble(
       colname = str_glue('`{.y}`'),
@@ -48,22 +48,24 @@ tbl_summary_1 <-
   select(age, grade, response, trt) %>%
   tbl_summary(by = trt)
 
-tbl_summary_1 %>%
+if (save_tables)
+  tbl_summary_1 %>%
   as_gt() %>%
   gtsave(file = "images/summary_basic.png")
 
-tibble::tribble(
-  ~Argument,       ~Description,
-  "`label=`",       "specify the variable labels printed in table",
-  "`type=`",        "specify the variable type (e.g. continuous, categorical, etc.)",
-  "`statistic=`",   "change the summary statistics presented",
-  "`digits=`",      "number of digits the summary statistics will be rounded to",
-  "`missing=`",     "whether to display a row with the number of missing observations",
-  "`missing_text=`","text label for the missing number row",
-  "`sort=`",        "change the sorting of categorical levels by frequency",
-  "`percent=`",     "print column, row, or cell percentages",
-  "`include=`",     "list of variables to include in summary table"
-) %>%
+if (save_tables)
+  tibble::tribble(
+    ~Argument,       ~Description,
+    "`label=`",       "specify the variable labels printed in table",
+    "`type=`",        "specify the variable type (e.g. continuous, categorical, etc.)",
[…]

### 2021-06-21 tex_tables/tbl_trial_desc.tex
formatting update

@@
 \texttt{trt} & Chemotherapy Treatment & character & \texttt{Drug A}, \texttt{Drug B} \\ 
-\texttt{age} & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17} , ... \\ 
-\texttt{marker} & Marker Level (ng/mL) & numeric & \texttt{0.003}, \texttt{0.005}, \texttt{0.013}, \texttt{0.015} , ... \\ 
+\texttt{age} & Age & numeric & \texttt{6}, \texttt{9}, \texttt{10}, \texttt{17}, ... \\ 
+\texttt{marker} & Marker Level (ng/mL) & numeric & \texttt{0.003}, \texttt{0.005}, \texttt{0.013}, \texttt{0.015}, ... \\ 
 \texttt{stage} & T Stage & factor & \texttt{T1}, \texttt{T2}, \texttt{T3}, \texttt{T4} \\ 
@@
 \texttt{death} & Patient Died & integer & \texttt{0}, \texttt{1} \\ 
-\texttt{ttdeath} & Months to Death/Censor & numeric & \texttt{3.53}, \texttt{5.33}, \texttt{6.32}, \texttt{7.27} , ... \\ 
+\texttt{ttdeath} & Months to Death/Censor & numeric & \texttt{3.53}, \texttt{5.33}, \texttt{6.32}, \texttt{7.27}, ... \\ 
 \bottomrule

### 2021-06-21 gtsummary_submission.R

@@ -25,11 +25,11 @@ if (save_tables)
       class = class(.x),
       values = trial %>% select(.y) %>% distinct() %>%
         arrange(!!rlang::sym(.y)) %>% slice(1:4) %>% pull(.y) %>%
-        na.omit() %>% {paste("`", ., "`", collapse = ", ")}
+        na.omit() %>% {paste0("`", ., "`", collapse = ", ")}
     ) %>%
       mutate(
         values = ifelse(length(unique(na.omit(trial[[.y]]))) > 4,
-                        paste(values, ", ..."),
+                        paste0(values, ", ..."),
                         values)
       )
   ) %>%

### 2021-06-21 gtsummary_submission.tex
image related updates
overleaf

@@
 The table merging and stacking abilities allows analysts to easily synthesize and compare output from several tables and share information in a compact format.
-All tables in this manuscript were created using \pkg{gtsummary} v1.4.0.
+All tables in this manuscript were created using \pkg{gtsummary} v1.4.1.
 
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{summary_basic.png}
+  \includegraphics[scale=0.49]{summary_basic.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{summary_plus.png}
+  \includegraphics[scale=0.49]{summary_plus.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{summary_plus_plus.png}
+  \includegraphics[scale=0.49]{summary_plus_plus.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{svysummary.png}
+  \includegraphics[scale=0.49]{svysummary.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{cross.png}
+  \includegraphics[scale=0.49]{cross.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{survfit.png}
+  \includegraphics[scale=0.49]{survfit.png}
   \centering
@@
 \begin{widefigure}[h!]
-  \includegraphics[scale=0.73]{custom.png}
+  \includegraphics[scale=0.49]{custom.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{regression.png}
+  \includegraphics[scale=0.49]{regression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{uvregression.png}
+  \includegraphics[scale=0.49]{uvregression.png}
   \centering
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{merge.png}
+  \includegraphics[scale=0.49]{merge.png}
   \centering
@@
 
-For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence internal is displayed in the same column as the model coefficients.
+For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence interval is displayed in the same column as the model coefficients.
 
@@
 \begin{figure}[h!]
-  \includegraphics[scale=0.73]{jama.png}
+  \includegraphics[scale=0.49]{jama.png}
   \centering

### 2021-07-15 gtsummary_submission.tex

@@
 The \pkg{gtsummary} package was created to streamline these everyday analysis tasks by allowing users to easily create reproducible summaries of data sets, regression models, survey data, and survival data with a simple interface and very little code.
-The package follows a tidy framework, making it easy to integrate in standard data workflows, and offers many table customization features through function arguments, helper functions, and custom themes.
+The package follows a tidy framework, making it easy to integrate with standard data workflows, and offers many table customization features through function arguments, helper functions, and custom themes.
 }
@@
 This package works to close the gap between a reproducible RMarkdown report and the final report.
-Specifically, \pkg{gtsummary} allows the us to fully customize and format summary tables with code, eliminating the need to modify any tables by hand after the table has been exported.
-Removing the need to modify tables after the table has been created eliminates an error-prone step in our workflow, and increases the reproducibility of our analyses and reports.
+Specifically, \pkg{gtsummary} allows the user to fully customize and format summary tables with code, eliminating the need to modify any tables by hand after the table has been exported.
+Removing the need to modify tables after the table has been created eliminates an error-prone step in our workflow and increases the reproducibility of our analyses and reports.
 
 Using \pkg{gtsummary}, analysts can easily summarize data frames, present and compare descriptive statistics between groups, summarize regression models, and report statistics inline in RMarkdown reports.
-After identifying these basic structures of most tables presented in medical literature (and other fields), we wrote \pkg{gtsummary} to ease creation of fully-formatted, ready to publish tables.
+After identifying these basic structures of most tables presented in the medical literature (and other fields), we wrote \pkg{gtsummary} to ease the creation of fully-formatted, ready-to-publish tables.
 
 Additionally, \pkg{gtsummary} leverages other analysis and tidying R packages to create a complete analysis and reporting framework.
-For example, we take advantage of the existing \CRANpkg{broom} tidiers to prepare regression results for \code{tbl\_regression()} and use \CRANpkg{gt} \citep{gt} to print \pkg{gtsummary} tables to various output formats (e.g.
-HTML, PDF, Word, or RTF).
+For example, we take advantage of the existing \CRANpkg{broom} \citep{broom} tidiers to prepare regression results for \code{tbl\_regression()} and use \CRANpkg{gt} \citep{gt} to print \pkg{gtsummary} tables to various output formats (e.g., HTML, PDF, Word, or RTF).
 Furthermore, \pkg{gtsummary} functions are designed to work within a "tidy" framework, utilizing the \CRANpkg{magrittr} \citep{magrittr} pipe operator and \CRANpkg{tidyselect} \citep{tidyselect} functions used throughout the \CRANpkg{tidyverse} \citep{tidyverse}.
 
-While other R packages are available to present data and regression model summary tables, such as \CRANpkg{skimr}, \CRANpkg{stargazer}, \CRANpkg{finalfit}, and \CRANpkg{tableone}, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize content of tables to a high degree.
+While other R packages are available to present data and regression model summary tables, such as \CRANpkg{skimr}, \CRANpkg{stargazer}, \CRANpkg{finalfit}, and \CRANpkg{tableone}, \pkg{gtsummary} is unique in that it is a one-stop-shop for most types of statistical tables and offers diverse features to customize the content of tables to a high degree.
 The default \CRANpkg{gtsummary} table is suitable to be published in a scientific journal with little or no additional formatting.
-For example, \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types so the user does not have to specify if a variable should be displayed with categorical or continuous summaries, which yields summary tables with minimal code.
+For example, \CRANpkg{gtsummary} has specific internal algorithms to identify variable data types, so there is no need for users to specify whether a variable should be displayed with categorical or continuous summaries, which yields summary tables with minimal code.
 
-Along with descriptive summaries, \pkg{gtsummary} summarizes statistical models, survey data, and builds cross tabulations.
+Along with descriptive summaries, \pkg{gtsummary} summarizes statistical models, survey data, survival data and builds cross-tabulations.
 After data are summarized in a table, \pkg{gtsummary} allows users to combine tables, either side-by-side (with \code{tbl\_merge()}) , or on top of each other (with \code{tbl\_stack()}).
@@
 To showcase \pkg{gtsummary} functions, we will use a simulated clinical trial data set containing baseline characteristics of 200 patients who received Drug A or Drug B, as well as the outcomes of tumor response and death.
-Each variable in the data frame has been assigned an attribute label with the \pkg{labelled} package \citep{labelled}, e.g.
-\code{trial \%>\% set\_variable\_labels(age = "Age")}, that will be shown in the summary tables.
+Each variable in the data frame has been assigned an attribute label with the \pkg{labelled} package \citep{labelled}, e.g., \code{trial \%>\% set\_variable\_labels(age = "Age")}, that will be shown in the summary tables.
 These labels are displayed in the \pkg{gtsummary} tables by default, and had labels not been assigned, the variable name would have been shown.
@@
 The \code{tbl\_summary()} function can take, at minimum, a data frame as the only input, and returns descriptive statistics for each column in the data frame.
-This is often the first table of clinical manuscripts and describes characteristics of the study cohort.
+This is often the first table of clinical manuscripts and describes the characteristics of the study cohort.
 A simple example is shown below.
@@
 The function is highly customizable, and it is initiated with sensible default settings.
-Specifically, \code{tbl\_summary} detects variable types of input data and calculates descriptive statistics accordingly.
+Specifically, \code{tbl\_summary()} detects variable types of input data and calculates descriptive statistics accordingly.
 For example, variables coded as \code{0/1}, \code{TRUE/FALSE}, and \code{Yes/No} are presented dichotomously.
-Additionally, \code{NA} values are recognized as missing and listed as unknown, and if a data set is labelled, the label attributes are utilized.
+Additionally, \code{NA} values are recognized as missing and listed as unknown, and if a data set is labeled, the label attributes are utilized.
 
@@
 For continuous variables, tables display one row of statistics per variable by default.
-This can be customized, and in the example below the age variable is cast to \code{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table.
+This can be customized, and in the example below, the age variable is cast to \code{"continuous2"} type, meaning the continuous summary statistics will appear on two or more rows in the table.
 This allows the number of non-missing observations and the mean to be displayed on separate lines.
@@
 \textbf{A note about notation:}
-Throughout the \pkg{gtsummary} package, you'll find function arguments that accept a list of formulas (or a single formula) as the input.
+Throughout the \pkg{gtsummary} package, you will find function arguments that accept a list of formulas (or a single formula) as the input.
 In the example above, the label for the age variable was updated using \code{label = age $\sim$ "Patient Age"}---equivalently, \code{label = list(age $\sim$ "Patient Age")}.
@@
 The \code{all\_continuous()} selector is a convenient way to select all continuous variables.
-In the example above it could have been used to change the summary statistics for all continuous variables---\code{all\_continuous() $\sim$ c("\{N\_nonmiss\}", "\{mean\} (\{sd\})")}.
-Similarly, users may utilize \code{all\_categorical()} (from \pkg{gtsummary}), or any of the \pkg{tidyselect} helpers used throughout the \pkg{tidyverse} package, such as \code{starts\_with()}, \code{contains()}, etc.
+In the example above, it could have been used to change the summary statistics for all continuous variables---\code{all\_continuous() $\sim$ c("\{N\_nonmiss\}", "\{mean\} (\{sd\})")}.
+Similarly, users may utilize \code{all\_categorical()} (from \pkg{gtsummary}) or any of the \pkg{tidyselect} helpers used throughout the \pkg{tidyverse} packages, such as \code{starts\_with()}, \code{contains()}, etc.
 
@@
 
-In the example below, descriptive statistics are shown by the treatment received and overall, as well as a p-value comparing the values between the treatments.
+In the example below, descriptive statistics are shown by the treatment received and overall, as well as a \emph{p}-value comparing the values between the treatments.
 Default statistical tests are chosen based on data type, and the statistical test performed can be customized in the \code{add\_p()} function.
-P-value formatting can be adjusted using the \code{pvalue\_fun=} argument which accepts both a proper function, as well the formula shortcut notation used throughout the \pkg{tidyverse} packages.
+\emph{p}-value formatting can be adjusted using the \code{pvalue\_fun=} argument, which accepts both a proper function, as well the formula shortcut notation used throughout the \pkg{tidyverse} packages.
 
@@
 
-The \code{tbl\_svysummary()} function is analogous to \code{tbl\_summary()} except a \CRANpkg{survey} \citep{survey} object is supplied rather than a data frame.
-The summary statistics presented take into account the survey weights, as do any p-values presented.
+The \code{tbl\_svysummary()} function is analogous to \code{tbl\_summary()}, except a \CRANpkg{survey} \citep{survey} object is supplied rather than a data frame.
+The summary statistics presented take into account the survey weights, as do any \emph{p}-values presented.
 
@@
 These functions work with any table constructed with \pkg{gtsummary}.
-Most common uses are changing the column headers and footnotes, or modifying the look of tables through bolding and italicization.
+The most common uses are changing the column headers and footnotes or modifying the look of tables through bolding and italicization.
 
@@
 
-Regression modelling is one of the most common tools of medical research.
+Regression modeling is one of the most common tools of medical research.
 The \pkg{gtsummary} package has two functions to help analysts prepare tabular summaries of regression models: \code{tbl\_regression()} and \code{tbl\_uvregression()}.
@@
 Like \code{tbl\_summary()}, \code{tbl\_regression()} creates highly customizable analytic tables with sensible defaults.
-Common regression models, such as logistic regression and Cox proportional hazards regression, are automatically identified and the tables headers are pre-filled with appropriate column headers (i.e.
-Odds Ratio and Hazard Ratio).
+Common regression models, such as logistic regression and Cox proportional hazards regression, are automatically identified, and the tables headers are pre-filled with appropriate column headers (i.e., Odds Ratio and Hazard Ratio).
 
 In the example below, the logistic regression model is summarized with \code{tbl\_regression()}.
-Note that a reference row for Grade has been added, and the variable labels have been carried through into the table.
+Note that a reference row for grade has been added, and the variable labels have been carried through into the table.
 Using \code{exponentiate = TRUE}, we exponentiate the regression coefficients, yielding the odds ratios.
-The helper function \code{add\_global\_p()} was used to replace the p-values for each term with the global p-value for grade.
+The helper function \code{add\_global\_p()} was used to replace the \emph{p}-values for each term with the global \emph{p}-value for grade.
 
@@
 The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} package's \code{tidy()} function to perform the initial formatting of the regression object \citep{broom}.
-Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice} and many more.
-A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals, or results with modified variance-covariance standard errors.
+Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.
+A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals or results with modified variance-covariance standard errors.
 
@@
 Imagine you need to report the results for age from the univariate table above.
-Typically, the odds ratio, confidence interval, and p-value would be hard-coded into a report, which can lead to reproducibility issues if the data is updated and the hard-coded statistics are not amended.
+Typically, the odds ratio, confidence interval, and \emph{p}-value would be hard-coded into a report, which can lead to reproducibility issues if the data is updated and the hard-coded statistics are not amended.
 A simple call to the \code{inline\_text()} function will dynamically add the model results to an RMarkdown report.
@@
 
-Here's how the line will appear in your report.
+Here is how the line will appear in your report.
 
@@
 This can be achieved by merging or stacking \pkg{gtsummary} tables using the \code{tbl\_merge()} and \code{tbl\_stack()} functions.
-For example, in cancer research we often report models predicting a tumor's response to treatment and risk of death.
-These two models are often reported side-by-side, and it's simple to construct using \code{tbl\_merge()}.
+For example, in cancer research we often report models predicting a tumor’s response to treatment and risk of death side-by-side in publications.
+This type of table is simple to construct using \code{tbl\_merge()}.
 First, build a table for each regression model using \code{tbl\_regression()}, then merge the two tables with \code{tbl\_merge()}.
@@
 We love themes.
-The default styling (e.g.
-statistics displayed in \code{tbl\_summary()}, how p-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
-However, you'll likely submit to another journal, or your personal preferences differ from the defaults.
+The default styling (e.g., statistics displayed in \code{tbl\_summary()}, how \emph{p}-values are rounded, decimal separator, and more) follow the reporting guidelines from European Urology, The Journal of Urology, Urology, and the British Journal of Urology International \citep{assel2019guidelines}.
+However, you will likely submit to another journal, or your personal preferences differ from the defaults.
 The \pkg{gtsummary} package is unique from other table building packages in the ability to set fine-grained customization defaults with themes.
 Themes were created to make these customizations easy to navigate and reuse across documents or projects.
-With themes, users can control default settings for existing functions (e.g.
-always present means instead of medians in \code{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
+With themes, users can control default settings for existing functions (e.g., always present means instead of medians in \code{tbl\_summary()}), as well as other changes that are not modifiable with function arguments.
 Several themes are available to follow various journals' reporting guidelines, reduce cell padding and font size, and language themes to translate \pkg{gtsummary} tables to more than 14 languages.
 
-For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large p-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence interval is displayed in the same column as the model coefficients.
+For example, using the theme for \textit{The Journal of the American Medical Association} (\textit{JAMA}), large \emph{p}-values are rounded to two decimal places, confidence intervals are shown as \code{"lb to ub"} instead of \code{"lb, ub"}, and the confidence interval is displayed in the same column as the model coefficients.
 
@@
 A custom theme was used to construct the \pkg{gtsummary} tables shown in this manuscript to match the \emph{R Journal} font and reduce the default cell padding.
-Themes are an evolving feature and we welcome additions of new journals' reporting guidelines or other themes useful to users.
+Themes are an evolving feature, and we welcome additions of new journals' reporting guidelines or other themes useful to users.
 A full glossary of customizable theme elements is available in the package's themes vignette (\url{http://www.danieldsjoberg.com/gtsummary/articles/themes.html}).
@@
 Tables printed with \pkg{gtsummary} can be seamlessly integrated into RMarkdown documents and knitted into various output types using a number of print engines.
-The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
-While \pkg{gt} is used as default for most outputs, you can easily use your print engine of choice with the conversion helper functions provided in the package (e.g. \code{as\_flex\_table()}).
-It's possible to get results in HTML, PDF (via \LaTeX), RTF, Microsoft Word, PowerPoint, Excel and others, utilizing the various print engines.
-The package is designed to interact with these print engines behind the scenes to reduce burden on users, and you generally only need to be aware of them if you want to add advanced customizations.
+The package was written to be a companion to the \pkg{gt} package from RStudio and is optimized to leverage the advanced customization features of this print engine, but offers compatibility with a variety of popular printing methods, including \code{knitr::kable()} \citep{knitr}, \CRANpkg{flextable}  \citep{flextable}, \CRANpkg{huxtable} \citep{huxtable}, and \CRANpkg{kableExtra} \citep{kableExtra}.
+While \pkg{gt} is used as the default for most outputs, you can easily use your print engine of choice with the conversion helper functions provided in the package (e.g., \code{as\_flex\_table()}).
+It is possible to get results in HTML, PDF (via \LaTeX), RTF, Microsoft Word, PowerPoint, Excel, and others, utilizing the various print engines.
+The package is designed to interact with these print engines behind the scenes to reduce the burden on users, and you generally only need to be aware of them if you want to add advanced customizations.
 
@@
 
-The functions in the \pkg{gtsummary} package were designed to reduce the burden of reporting, and to work together to easily construct both simple and complex tables.
+The functions in the \pkg{gtsummary} package were designed to reduce the burden of reporting and to work together to easily construct both simple and complex tables.
 It is our hope that the user-friendly syntax and publication-ready tables will aid analysts in preparing reproducible and high-quality findings.

### 2021-07-16 gtsummary_submission.R
tex table caption udpates

@@ -12,7 +12,7 @@ set_gtsummary_theme(my_r_journal_theme)
 
 # set to TRUE to re-render all tables (gt images and latex tables)
 # if you're at the R Journal, please leave this as FALSE
-save_tables <- FALSE
+save_tables <- TRUE
 
 # Data Summaries ---------------------------------------------------------------
 
@@ -38,7 +38,7 @@ if (save_tables)
   cols_align("left", everything()) %>%
   as_latex() %>%
   as.character() %>%
-  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}Example data frame, \\texttt{trial}}\\\\\n\\toprule")) %>%
+  str_replace(fixed("\\bottomrule"), fixed("\\bottomrule\\caption{\\label{tab:caption}Table 1. Example data frame, \\texttt{trial}}\\\\\n")) %>%
   readr::write_lines("tex_tables/tbl_trial_desc.tex")
 
 ## `tbl_summary()` -------------------------------------------------------------
@@ -57,7 +57,7 @@ if (save_tables)
   tibble::tribble(
     ~Argument,       ~Description,
     "`label=`",       "specify the variable labels printed in table",
-    "`type=`",        "specify the variable type (e.g. continuous, categorical, etc.)",
+    "`type=`",        "specify the variable type (e.g., continuous, categorical, etc.)",
     "`statistic=`",   "change the summary statistics presented",
     "`digits=`",      "number of digits the summary statistics will be rounded to",
     "`missing=`",     "whether to display a row with the number of missing observations",
@@ -70,7 +70,7 @@ if (save_tables)
   fmt_markdown(everything()) %>%
   as_latex() %>%
   as.character() %>%
-  str_replace(fixed("\\toprule"), fixed("\\caption{\\label{tab:}\\texttt{tbl\\_summary()} function arguments}\\\\\n\\toprule")) %>%
+  str_replace(fixed("\\bottomrule"), fixed("\\bottomrule\\caption{\\label{tab:caption}Table 2. \\texttt{tbl\\_summary()} function arguments}\\\\\n")) %>%
   readr::write_lines("tex_tables/tbl_tbl_summary_args.tex")
 
 tbl_summary_2 <-
@@ -94,20 +94,19 @@ if (save_tables)
 if (save_tables)
   tibble::tribble(
     ~Function,             ~Description,
-    "`add_p()`",           "add p-values to the output comparing values across groups",
+    "`add_p()`",           "add *p*-values to the output comparing values across groups",
[…]

### 2021-07-16 gtsummary_submission.R
Update gtsummary_submission.R

@@ -12,7 +12,7 @@ set_gtsummary_theme(my_r_journal_theme)
 
 # set to TRUE to re-render all tables (gt images and latex tables)
 # if you're at the R Journal, please leave this as FALSE
-save_tables <- TRUE
+save_tables <- FALSE
 
 # Data Summaries ---------------------------------------------------------------
 

### 2021-07-16 gtsummary_submission.tex
Updates from Overleaf
overleaf

@@
 \title{Reproducible Summary Tables with the \pkg{gtsummary} Package}
-\author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry}
+\author{by Daniel D. Sjoberg, Karissa Whiting, Michael Curry, Jessica A. Lavery, Joseph Larmarange}
 
@@
 
-The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} package's \code{tidy()} function to perform the initial formatting of the regression object \citep{broom}.
+The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom}\citep{broom.helpers}.
 Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.
@@
   \email{currym1@mskcc.org}}
+
+ \address{Jessica A. Lavery\\
+  Memorial Sloan Kettering Cancer Center\\
+  1275 York Ave., New York, New York 10022\\
+  USA\\
+  ORCiD: 0000-0002-2746-5647\\
+  \email{laveryj@mskcc.org}}
+
+\address{Joseph Larmarange\\
+  Centre Population \& Développement, IRD, Université de Paris, Inserm\\
+  45 rue des Saints-Pères 75006 PARIS\\
+  France\\
+  ORCID: 0000-0001-7097-700X\\
+  \email{joseph.larmarange@ceped.org}}

### 2021-07-16 gtsummary_submission.tex
overleaf

@@
 
-The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom}\citep{broom.helpers}.
+The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom,broom.helpers}.
 Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.

### 2021-07-16 gtsummary_submission.tex
Updates from Overleaf

@@
 
-The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom,broom.helpers}.
+The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom}, \CRANpkg{parameters}, and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom,broom.helpers}.
 Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.

### 2021-07-17 gtsummary_submission.tex
overleaf

--- 2021-07-16 gtsummary_submission.tex
@@
 
-The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom} and \CRANpkg{broom.helpers} packages to perform the initial formatting of the regression object \citep{broom,broom.helpers}.
-Because \code{tbl\_regression()} utilizes \code{broom::tidy()}, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.
+The \code{tbl\_regression()} function leverages the huge effort behind the \pkg{broom}, \CRANpkg{parameters} \citep{parameters}, and \CRANpkg{broom.helpers} \citep{broom.helpers} packages to perform the initial formatting of the regression object.
+Because \code{tbl\_regression()} utilizes these packages, there are many model types that are supported out of the box, such as \code{lm()}, \code{glm()}, \code{lme4::lmer()}, \code{lme4::glmer()}, \code{geepack::geeglm()}, \code{survival::coxph()}, \code{survival::survreg()}, \code{survival::clogit()}, \code{nnet::multinom()}, \code{rstanarm::stan\_glm()}, models built with the \pkg{mice} package \citep{mice}, and many more.
 A custom tidier may be specified as well, which is helpful when you need to present non-standard modifications to your model results such as Wald confidence intervals or results with modified variance-covariance standard errors.
@@
 However, you will likely submit to another journal, or your personal preferences differ from the defaults.
-The \pkg{gtsummary} package is unique from other table building packages in the ability to set fine-grained customization defaults with themes.
+The \pkg{gtsummary} package is unique from other table building packages with the ability to set fine-grained customization defaults with themes.
 Themes were created to make these customizations easy to navigate and reuse across documents or projects.
@@
 \end{figure}
+
+The language theme can be used to translate the table to another language and allows users to specify the decimal and big mark symbols.
+For example, \code{theme\_gtsummary\_language(language = "es", decimal.mark = ",", big.mark = ".")} will translate the output to Spanish and format numeric results as \code{1.000,00} instead of \code{1,000.00} (the default formatting).
 
